---
title: "Homework 1"
author: "Tomasz Zaremba"
date: "2015-10-08"
output: 
  html_document:
    toc: TRUE
---

# Home work

Use `knitr` to create a report with numerical and graphical summaries for GermanCredit.

Summarize following variables: Duration, Amount, Class, Personal.Male.Single.

```{r, cache=TRUE}

# Load the set
library(caret)
data(GermanCredit)

```

## Example values from columns "Duration", "Amount", "Class" and "Personal.Male.Single":

```{r, cache=TRUE}
# Let's see what our columns look like
head(GermanCredit[, c("Duration", "Amount", "Class", "Personal.Male.Single")], 10)
```

## Numerical summaries for single variables:

```{r, cache=TRUE}
# Contingency tables for discrete variables
table(GermanCredit$Class)
table(GermanCredit$Personal.Male.Single)

# Contingency tables for continuous variables (in intervals)
table(cut(GermanCredit$Duration, 4))
table(cut(GermanCredit$Amount, 4))

# Summaries for continuous variables
summary(GermanCredit$Duration)
summary(GermanCredit$Amount)
```

## Graphical summaries for single variables.

Barplots:

```{r, cache=TRUE}

# The majority of credits were paid
classTable = table(GermanCredit$Class)
barplot(classTable)

# Most single males paid the credit
singleMaleTable = table(GermanCredit$Personal.Male.Single)
barplot(singleMaleTable)

amountTable = table(cut(GermanCredit$Amount, 5))
barplot(amountTable, horiz=TRUE, las=1)

durationTable = table(cut(GermanCredit$Duration, 5))
barplot(durationTable, horiz=TRUE, las=1)

```

Histograms:

```{r, cache=TRUE}
# Credits were rather short
hist(GermanCredit$Duration, col="lightblue")
# And small...
hist(GermanCredit$Amount, 50, col="lightgreen")
```

Boxplots:

```{r, cache=TRUE}
# Boxplots confirm what we observed on the histograms
boxplot(GermanCredit$Duration, horizontal=TRUE)
boxplot(GermanCredit$Amount, horizontal=TRUE)
```

## Numerical summaries for pairs of variables.

Proportions tables:
```{r, cache=TRUE}
singleMaleClassTable = table(GermanCredit$Class, GermanCredit$Personal.Male.Single)
singleMaleClassTable
prop.table(singleMaleClassTable, 1)

amountClassTable = table(GermanCredit$Class, cut(GermanCredit$Amount, 6))
amountClassTable
round(100*prop.table(amountClassTable, 1), 1)

durationClassTable = table(GermanCredit$Class, cut(GermanCredit$Duration, 6))
durationClassTable
round(100*prop.table(durationClassTable, 1), 1)
```

```{r, cache=TRUE}
# Quantiles for amount in two classes: Bad and Good
by(GermanCredit$Amount, GermanCredit$Class, summary)

# Quantiles for duration in two classes: Bad and Good
by(GermanCredit$Duration, GermanCredit$Class, summary)
```

## Graphical summaries for pairs of variables.

Moisaic plots showing relations between two variables:

```{r, cache=TRUE}
# Most single males paid the credit:
singleMaleClassTable = table(GermanCredit$Class, GermanCredit$Personal.Male.Single)
mosaicplot(singleMaleClassTable, main="Single males vs. Class", color = c("lightblue", "lightgreen"))

durationClassTable = table(GermanCredit$Class, cut(GermanCredit$Duration, c(0, 9, 15, 24), c("Short", "Average", "Long")))
mosaicplot(durationClassTable, main="Duration vs. Class", color = c("lightblue", "lightgreen"))

amountClassTable = table(GermanCredit$Class, cut(GermanCredit$Amount, c(0, 5000, 10000, 20000), c("Small", "Average", "Big")))
mosaicplot(amountClassTable, main="Credit amount vs. Class", color = c("lightblue", "lightgreen"))
```

Boxplots for continuous variables:

```{r, cache=TRUE}
# Distribution of amount in classes Good and Bad
boxplot(GermanCredit$Amount ~ GermanCredit$Class, horizontal = TRUE)
# Distribution of Duration in classes Good and Bad
boxplot(GermanCredit$Duration ~ GermanCredit$Class, horizontal = TRUE)
```

Violin plot combining a boxplot and a kernel density plot:

```{r, cache=TRUE}
library(vioplot)
par(mar=c(5,5,2,2))
# Relation between Duration and Amount - the majority of credits over 5000 where for ~ 25-40 months
vioplot(GermanCredit$Duration[GermanCredit$Amount < 5000],
        GermanCredit$Duration[GermanCredit$Amount >= 5000])
```
