---
title: "Phase 1 - Project 2"
author: "Tomasz K, Tomasz Z, Margareta K"
date: "2015-12-09"
output: 
  html_document:
    self_contained: no
    toc: TRUE
---

# Project phase description
In the first phase you should characterize types of interactions with different machines (exploratory analysis). Try to answer following questions:

- what is the average (or distribution) time of interaction with a given machine
- is there a difference in the time of interaction and weekday or hour
- what is the average (or distribution) length of playing in the space with stations

# Loading data
```{r, warning=FALSE, message=FALSE, cache=TRUE}
load("C:\\Users\\Tomek\\Desktop\\Studia\\MSc\\Semestr 2\\Data mining\\Projekty\\4\\verySmallLogs.rda")
library(lubridate)
verySmallLogs$dayOfWeek = strftime(verySmallLogs$date,'%A')
verySmallLogs$dayOfWeekNumber = as.numeric(strftime(verySmallLogs$date,'%u'))
verySmallLogs$hour = substr(as.character(verySmallLogs$date), 12, 13);
verySmallLogs = verySmallLogs[order(verySmallLogs$dayOfWeekNumber), ]
verySmallLogs = verySmallLogs[!verySmallLogs$visitor == -1, ]
```

# General statistics for machines

## Average times spent on a machine

```{r, warning=FALSE, message=FALSE, cache=TRUE}
library(ggplot2)
library(sqldf)

time_per_visitor <- sqldf('SELECT station, visitor, MAX(date) - MIN(date) as time FROM verySmallLogs GROUP BY station, visitor')
ggplot(data=time_per_visitor,aes(x=time))+geom_bar(binwidth=1)+xlim(0,1000)
```

## Time spent by visitors in stations
```{r, warning=FALSE, message=FALSE, cache=TRUE}
time_per_visitor = time_per_visitor[time_per_visitor$time < 500, ]

# Distribution of time per visitor in a station
summary(time_per_visitor$time)

# Boxplot of time per visitor in a station
boxplot(time_per_visitor$time, horizontal = TRUE)
```

We can see that the vast majority of interactions last less than 200 seconds. Even though we eliminated the biggest outliers, the mean is not the best criterion to assess the time spent by visitors at each station. Because of this, we will use median in our further calculations.

## Median time spent on each machine

```{r, warning=FALSE, message=FALSE, cache=TRUE, fig.width = 15, fig.height=8}
machineStatistics <- sqldf('SELECT station, MEDIAN(time) as median_time, count(*) as numberOfVisitors FROM time_per_visitor GROUP BY station')
machineStatistics

barplot(machineStatistics$median_time, main="Median time on each machine", 
  	xlab="Machine", names.arg = machineStatistics$station)
```

We can see that the biggest amount of time on average was spent on cnk38, the smallest on cnk19a.

## Number of visitors on each machine

```{r, warning=FALSE, message=FALSE, cache=TRUE, fig.width = 15, fig.height=8}

barplot(machineStatistics$numberOfVisitors, main="Number of visitors on each machine", 
  	xlab="Machine", names.arg = machineStatistics$station)

```

We can see that cnk19a was the most popular machine and cnk38 was the least popular one. This seems logical as these are the two machines which had the smallest and the biggest average time of interaction respectively. We can suspect that the small number of visitors using cnk38 comes from the fact that on average it was used the longest. The great number of visitors using cnk19a may come from the fact that on average it was used for a short time so many people could try it.

# Statistics for machines on particular days and hours

```{r, warning=FALSE, message=FALSE, cache=TRUE}
time_per_day_visitor = sqldf('SELECT station, visitor, MAX(date) - MIN(date) as time, dayOfWeek, dayOfWeekNumber FROM verySmallLogs GROUP BY station, visitor, dayOfWeekNumber')
time_per_day_visitor = time_per_day_visitor[time_per_day_visitor$time < 500, ]

# Median time and the number of visitors per machine and weekday
machineStatisticsWeekday <- sqldf('SELECT station, dayOfWeek, dayOfWeekNumber, AVG(time) as median_time, count(*) as numberOfVisitors FROM time_per_day_visitor GROUP BY station, dayOfWeekNumber')

# Average time and the number of visitors per weekday
weekdayStatistics = sqldf('SELECT dayOfWeek, dayOfWeekNumber, AVG(median_time) as average_time_per_day, sum(numberOfVisitors) as numberOfVisitorsPerDay FROM machineStatisticsWeekday GROUP BY dayOfWeekNumber')

time_per_day_and_hour_visitor = sqldf('SELECT station, visitor, MAX(date) - MIN(date) as time, dayOfWeek, dayOfWeekNumber, hour FROM verySmallLogs GROUP BY station, visitor, dayOfWeekNumber, hour')
time_per_day_and_hour_visitor = time_per_day_and_hour_visitor[time_per_day_and_hour_visitor$time < 500, ]

# Median time and the number of visitors per machine, weekday and hour
machineStatisticsWeekdayHour <- sqldf('SELECT station, dayOfWeek, dayOfWeekNumber, hour, AVG(time) as median_time, count(*) as numberOfVisitors FROM time_per_day_and_hour_visitor GROUP BY station, dayOfWeekNumber, hour')

# Average time and the number of visitors per weekday and hour
weekdayHourStatistics = sqldf('SELECT dayOfWeek, dayOfWeekNumber, hour, AVG(median_time) as average_time_per_day_hour, sum(numberOfVisitors) as numberOfVisitors FROM machineStatisticsWeekdayHour GROUP BY dayOfWeekNumber, hour')

# Average time and the number of visitors per hour
hourStatistics = sqldf('SELECT hour, AVG(average_time_per_day_hour) as average_time_per_hour, sum(numberOfVisitors) as numberOfVisitorsPerHour FROM weekdayHourStatistics GROUP BY hour')

# Statistics per day and machine
machineStatisticsWeekday = machineStatisticsWeekday[, !(colnames(machineStatisticsWeekday) %in% c("dayOfWeekNumber"))]

# Statistics per day
weekdayStatistics = weekdayStatistics[, !(colnames(weekdayStatistics) %in% c("dayOfWeekNumber"))]

time_per_day_and_hour_visitor = time_per_day_and_hour_visitor[order(time_per_day_and_hour_visitor$dayOfWeekNumber), ]

# Statistics per day, hour and machine
machineStatisticsWeekdayHour = machineStatisticsWeekdayHour[, !(colnames(machineStatisticsWeekdayHour) %in% c("dayOfWeekNumber"))]

# Statistics per day, hour
weekdayHourStatistics = weekdayHourStatistics[, !(colnames(weekdayHourStatistics) %in% c("dayOfWeekNumber"))]
```

## Time spent by visitors in particular weekdays
```{r, warning=FALSE, message=FALSE, cache=TRUE, fig.width = 15, fig.height=8}

weekdayStatistics

# Distribution of time per weekday
summary(weekdayStatistics$average_time_per_day)

# Boxplot of time per weekday
boxplot(weekdayStatistics$average_time_per_day, horizontal = TRUE)

# Barplot of the time per weekday
barplot(weekdayStatistics$average_time_per_day, main="Average time per weekday", 
   xlab="Hours in a day", names.arg = weekdayStatistics$dayOfWeek)

```

We can observe that slightly more time is spent on average during the weekends.

## The number of visitors in particular weekdays
```{r, warning=FALSE, message=FALSE, cache=TRUE, fig.width = 15, fig.height=8}

weekdayStatistics

# Distribution of the number of visitors in particular weekdays
summary(weekdayStatistics$numberOfVisitorsPerDay)

# Boxplot of the number of visitors in particular weekdays
boxplot(weekdayStatistics$numberOfVisitorsPerDay, horizontal = TRUE)

# Barplot of the number of visitors in particular weekdays
barplot(weekdayStatistics$numberOfVisitorsPerDay, main="The number of visitors per day", 
   xlab="Hours in a day", names.arg = weekdayStatistics$dayOfWeek)
```

We can easily observe that the centre is closed on Mondays. The number of visitors is not 0 though. We might guess that some routine tests are performed then.

## Time spent by visitors on machines in particular weekdays
```{r, warning=FALSE, message=FALSE, cache=TRUE}

machineStatisticsWeekday

# Distribution of time per station and weekday
summary(machineStatisticsWeekday$median_time)

# Boxplot of time per station and weekday
boxplot(machineStatisticsWeekday$median_time, horizontal = TRUE)

```

## The number of visitors using machines in particular weekdays
```{r, warning=FALSE, message=FALSE, cache=TRUE}

# Distribution of the number of visitors per station and weekday
summary(machineStatisticsWeekday$numberOfVisitors)

# Boxplot of the number of visitors per station and weekday
boxplot(machineStatisticsWeekday$numberOfVisitors, horizontal = TRUE)

```

## Time spent by visitors on particular weekdays and hours

```{r, warning=FALSE, message=FALSE, cache=TRUE}

weekdayHourStatistics

# Distribution of the time per weekday and hour
summary(weekdayHourStatistics$average_time_per_day_hour)

# Boxplot of the time per weekday and hour
boxplot(weekdayHourStatistics$average_time_per_day_hour, horizontal = TRUE)

```

## The number of visitors on particular weekdays and hours

```{r, warning=FALSE, message=FALSE, cache=TRUE}

# Distribution of the number of visitors on particular weekdays and hours
summary(weekdayHourStatistics$average_time_per_day_hour)

# Boxplot of the number of visitors on particular weekdays and hours
boxplot(weekdayHourStatistics$average_time_per_day_hour, horizontal = TRUE)

```

## Time spent by visitors on machines for particular weekdays and hours
```{r, warning=FALSE, message=FALSE, cache=TRUE}

# Distribution of the time per station, weekday and hour
summary(machineStatisticsWeekdayHour$median_time)

# Boxplot of the time per station, weekday and hour
boxplot(machineStatisticsWeekdayHour$median_time, horizontal = TRUE)

```

## The number of visitors using machines on particular weekdays and hours
```{r, warning=FALSE, message=FALSE, cache=TRUE}

# Distribution of the number of visitors using machines on particular weekdays and hours
summary(machineStatisticsWeekdayHour$median_time)

# Boxplot of the number of visitors using machines on particular weekdays and hours
boxplot(machineStatisticsWeekdayHour$median_time, horizontal = TRUE)

```

## Time spent by visitors on particular hours

```{r, warning=FALSE, message=FALSE, cache=TRUE}

hourStatistics

# Distribution of the time per hour
summary(hourStatistics$average_time_per_hour)

# Boxplot of the time per hour
boxplot(hourStatistics$average_time_per_hour, horizontal = TRUE)

# Barplot of the average times per hour
barplot(hourStatistics$average_time_per_hour, main="Average times per hour", 
   xlab="Hours in a day", names.arg = hourStatistics$hour)

```

## The number of visitors on particular hours

```{r, warning=FALSE, message=FALSE, cache=TRUE}

# Distribution of the number of visitors per hour
summary(hourStatistics$numberOfVisitorsPerHour)

# Boxplot of the number of visitors per hour
boxplot(hourStatistics$numberOfVisitorsPerHour, horizontal = TRUE)

# Barplot of the number of visitors per hour
barplot(hourStatistics$numberOfVisitorsPerHour, main="The number of visitors per hour", 
   xlab="Hours in a day", names.arg = hourStatistics$hour)

```

We can observe that the number of visitors is the smallest in the morning and in the evening and the biggest durring the day, especially in the afternoon.

## Median times spent on each machine on each day

## The number of visitors using each machine on each day

Days of the week do not influence significantly the average time spent on a machine and the number of visitors using the machine. The machines 'cnk19a' and 'cnk38' stand out of the crowd every day.

# Conclusions

The majority of interactions last less than 200 seconds. There are few observations (outliers) with great times of interaction which affect the mean so the median is a better criterion to use when analyzing this data.

The centre is closed every day between 10 PM and 7 AM and is closed completely on Mondays. There are, however, some technical works performed then, so the number of registered visitors is not zero. We can see that 'cnk19a' is the most popular machine and 'cnk38' is the least popular one. On the other hand, the time spent on 'cnk19a' is the shortest and the time spent on 'cnk38' is on average the longest. We might suspect that 'cnk38' is usually used for a long time so people are blocking it. The machine 'cnk19a' is tried very often but only for a short time - maybe it's boring? If so, we could buy some more copies of the most popular machines, especially 'cnk38' so that more people can use it.

The number of visitors is the smallest in the morning and in the evening and the biggest durring the day, especially in the afternoon. The number of visitors does not differ significantly between days (except Monday, where the centre is closed).