---
title: "Project 2 - First Phase"
author: "Ziad Al Bkhetan, Neven Piculjan, Naveen Mupparapu"
date: "December 10, 2015"
output: 
  html_document:
    toc : TRUE
---

# Internet of Things - Introduction

In the first phase we should characterize types of interactions with different machines (exploratory analysis). We tried to answer following questions:

* What is the average (or distribution) time of interaction with a given machine
* Is there a relation between the duration of an interaction and variables like weekday or hour
* Other statistics

We Assumed That the visit is the duration between Enterance and Leaving Activity for the same visitor in the same station.

We prepared a dataset full of all needed information, then we used grouping function to group the data as we want.

# Data Loading And Preparation
In this step we will prepare the data set to start analysis phase, we assumed that the maybe the dataset is not sorted, so we sort it based on the visitor and date, which makes stations interactions detection easier.
<br> in this report we worked in a smaller dataset to make the performance faster.

```{r, warning=FALSE, message=FALSE}
library(tidyr)
library(dplyr)
library(ggplot2)

load('D:/MSc Computer science and informatics/3rd semester/Data Mining - Advances/Lab Projects/Project 2/newSmallLogs.rda')

# Sort the data based on Visitors and Date
index <- with(mydata, order(visitor, date))
orderedData <-mydata[index, ]

```

# Prepare The Final DataSet
in this step we will fetch all needed data for analysis and visualization, and the most important for us are: Visitor ID, Station, Consumed Time, Day, and Hour.
<br> we loop over the data set and try to find the visits between Entrance and Leaving for the same visitor and same station.
```{r, warning=FALSE, message=FALSE}

stations_time_lst  <- c()
stations_count_lst <- c()
stations_visit_time_lst <- c()
matRes <- NULL

for (i in 2:nrow(orderedData))
{
  if ((orderedData[i, 3] == orderedData[i - 1, 3]) & (as.character(orderedData[i, 2]) == as.character(orderedData[i - 1, 2])) & (as.character(orderedData[i, 4]) == 'Leaving') & (as.character(orderedData[i - 1, 4]) == 'Entering'))
  {
    if (length(stations_count_lst) == 0)
    {
      tim <- difftime(orderedData[i, 1], orderedData[i - 1, 1],  units = "secs")
      stations_count_lst[as.character(orderedData[i, 2])] <- 1
      stations_time_lst [as.character(orderedData[i, 2])] <- tim
      matRes <- matrix(nrow=1, ncol=5)
      colnames(matRes) <- c('Station', 'Visitor', 'Cons_time', 'WDay', 'THour')
      matRes[1, 1] = as.character(orderedData[i, 2])
      matRes[1, 2] = orderedData[i, 3]
      matRes[1, 3] = as.numeric(tim)
      matRes[1, 4] = as.POSIXlt(orderedData[i - 1, 1])$wday
      matRes[1, 5] = as.POSIXlt(orderedData[i - 1, 1])$hour
    }
    else if (is.na(stations_count_lst[as.character(orderedData[i, 2])]))
        {
          tim <- difftime(orderedData[i, 1], orderedData[i - 1, 1],  units = "secs")
          stations_count_lst[as.character(orderedData[i, 2])] <- 1
          stations_time_lst [as.character(orderedData[i, 2])] <- tim
          matRes <- rbind(matRes, c(as.character(orderedData[i, 2]), orderedData[i, 3], as.numeric(tim), as.POSIXlt(orderedData[i - 1, 1])$wday, as.POSIXlt(orderedData[i - 1, 1])$hour ))
        } 
        else
        {
          tim <- difftime(orderedData[i, 1], orderedData[i - 1, 1],  units = "secs")
          stations_count_lst[as.character(orderedData[i, 2])] <- stations_count_lst[as.character(orderedData[i, 2])] + 1
          stations_time_lst [as.character(orderedData[i, 2])] <- stations_time_lst [as.character(orderedData[i, 2])] + tim
          matRes <- rbind(matRes, c(as.character(orderedData[i, 2]), orderedData[i, 3], as.numeric(tim), as.POSIXlt(orderedData[i - 1, 1])$wday, as.POSIXlt(orderedData[i - 1, 1])$hour ))
        }
  }
}

finalData <- as.data.frame(matRes)
finalData$Cons_time <- as.numeric(finalData$Cons_time)

```

# Results Analysis
We Tried in this step to visualize all possible relations and plots which describe the data as much as possible, and give the user some hidded information.

## General Information
we will show the general information we got in this analysis:

```{r, warning=FALSE, message=FALSE}
cat("Visitors Count:", length(unique(finalData$Visitor)))
cat("Stations Count:", length(unique(finalData$Station)))
summary(finalData$Station)
print("Days Information:")
summary(finalData$WDay)
print("Hours Information:")
summary(finalData$THour)
```

Some Charts to make it clear

```{r, warning=FALSE, message=FALSE}
WeeksDays <- c("Sun", "Mon", "Tue", "Wed", "Thu", "Fri", "Sat")

station_count = finalData %>% 
  group_by(Station) %>%
  summarise(
    s_count = n()
  )     
x = station_count$s_count
names(x) <- station_count$Station
barplot(x, las=2)

visitor_count = finalData %>% 
  group_by(Visitor) %>%
  summarise(
    s_count = n()
  )     
x = visitor_count$s_count
names(x) <- visitor_count$Visitor
barplot(x, las=2) 


days_count = finalData %>% 
  group_by(WDay) %>%
  summarise(
    s_count = n()
  )     
x = days_count$s_count
names(x) <- WeeksDays[days_count$WDay]
barplot(x, las=2) 

hours_count = finalData %>% 
  group_by(THour) %>%
  summarise(
    s_count = n()
  )     
x = hours_count$s_count
names(x) <- hours_count$THour
barplot(x, las=2) 



```



## Consumed Time During Stations Interactions 

in this plot we will see the consumed time for each visitor in specific visit and station.

```{r, warning=FALSE, message=FALSE}
ggplot(finalData, aes(Visitor, Cons_time, color=Station)) +
  geom_point() +
  theme_bw() + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))

station_count = finalData %>% 
  group_by(Station) %>%
  summarise(
    s_count = n(),
    s_sum = sum(Cons_time),
    s_avg = mean(Cons_time)
  )     
station_count
```

## Consumed time In Each Day
this plot shows the average consumed time in each day in this dataset  
```{r, warning=FALSE, message=FALSE}
ggplot(finalData, aes(WDay, Cons_time, color=Station)) +
  geom_point() +
  theme_bw() + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))

days_count = finalData %>% 
  group_by(WDay) %>%
  summarise(
    s_count = n(),
    s_sum = sum(Cons_time),
    s_avg = mean(Cons_time)
  )     
x = days_count$s_avg
names(x) <- WeeksDays[days_count$WDay]
barplot(x, las=2) 

```

## Average Consumed time In Each Hour
this plot shows the average consumed time in each Hour in this dataset  

```{r, warning=FALSE, message=FALSE}
ggplot(finalData, aes(THour, Cons_time, color=Station)) +
  geom_point() +
  theme_bw() + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))

hours_count = finalData %>% 
  group_by(THour) %>%
  summarise(
    s_count = n(),
    s_sum = sum(Cons_time),
    s_avg = mean(Cons_time)
  )     
x = hours_count$s_avg
names(x) <- hours_count$THour
barplot(x, las=2) 

```

## Distinct Visitors In Each Day
this plot shows the count of distinct visitors in each day in this dataset  

```{r, warning=FALSE, message=FALSE}

visitor_count = finalData %>% 
  group_by(WDay) %>%
  summarise(
    s_count = length(unique(Visitor))
  )     
x = visitor_count$s_count
names(x) <- WeeksDays[visitor_count$WDay]
barplot(x, las=2) 

```

## Distinct Visitors In Each Hour
The same as the previous plot but in each hour 
```{r, warning=FALSE, message=FALSE}

visitor_count = finalData %>% 
  group_by(THour) %>%
  summarise(
    s_count = length(unique(Visitor))
  )     
x = visitor_count$s_count
names(x) <- visitor_count$THour
barplot(x, las=2) 

```

## Average Consumed Time in Each Hour in Each Day
Here we try to go more deep in these statistics , so we tried to find the consumed time for each hour in each day
```{r, warning=FALSE, message=FALSE}

time_d_h = finalData %>% 
  group_by(WDay, THour) %>%
  summarise(
    s_count = n(),
    s_sum = sum(Cons_time),
    s_avg = mean(Cons_time)
  )
lst <- unique(time_d_h$WDay)
for (d in 1:length(lst) )
{
  temp <- time_d_h[time_d_h$WDay == lst[d],]
  x = temp$s_avg
  names(x) <- paste(WeeksDays[temp$WDay], '_', temp$THour) 
  barplot(x, las=2) 
  
}

```

## Distinct Visitors in Each Hour in Each Day
The same as the previous plot, but for Distinct visitors
```{r, warning=FALSE, message=FALSE}

visitor_d_h = finalData %>% 
  group_by(WDay, THour) %>%
  summarise(
    s_count = length(unique(Visitor))
  ) 
lst <- unique(visitor_d_h$WDay)
for (d in 1:length(lst) )
{
  temp <- visitor_d_h[visitor_d_h$WDay == lst[d],]
  x = temp$s_count
  names(x) <- paste(WeeksDays[temp$WDay], '_', temp$THour) 
  barplot(x, las=2) 
  
}


```


# Remarks and Conclusion
In this task we applied this strategy, findg all needed data in one dataset, using dplyr library to group the data and get retrieve all needed information to visualize, that makes our task easier.
<br> for the results, we believe these plots show the most important information we got from this dataset
