---
title: "DMA - Homework 2"
author: "Viet Ba Mai"
date: "14 pazdziernika 2016"
output: 
  html_document:
  toc: TRUE
---

#Loading data
Firstly we need to load the data from a given .csv file `student-mat`.
```{r, cache=TRUE}
student_mat <- read.csv(file="C:/Users/vietba/Downloads/student-mat.csv", header=TRUE, sep=";", encoding="UTF-8")

```


#Input data
Below 10 first observations of the imported table are presented.

```{r, echo=FALSE, cache=TRUE}
head(student_mat, 10)
```

I chose the `Dalc` column as the target class. This column stands for workday alcohol consumption.
However factor needs to be applied on the `Dalc` variable in order to use it as the ideal value.

Below are 10 first observations of columns selected for the classifier and the factored response variable:

```{r, cache=TRUE}
#install.packages("caret")
library(caret)
studentnew <- within(student_mat, Dalc <- factor(Dalc))
head(studentnew[, c("Dalc", "studytime", "age", "absences", "failures", "freetime", "goout")], 10)
```

I chose these particular variables, because after observing the table it seems that they affect the value of `Dalc`. In other words they change with that value fairly proportionally.


#Preparing training and testing set
We divide the `student_mat` set into training and testing set with their ratio being 3:1 respectively.
```{r, cache=TRUE}
set.seed(1313)
indxTrain <- createDataPartition(y = studentnew$Dalc, p = 0.75)
str(indxTrain)

stdntmatTrain <- studentnew[indxTrain$Resample1,]
stdntmatTest <- studentnew[-indxTrain$Resample1,]
```


#Train and test
We build the classifier for selected variables and then train it on the training set `stdntmatTrain`.
```{r, cache=TRUE}
knnFit <- knn3(Dalc ~ studytime + age + absences + failures + freetime + goout, data = stdntmatTrain, k=20, prob=TRUE)
```
Above I performed training on 20 nearest neighbours model.


Testing on the testing set `stdntmatTest`.
```{r, cache=TRUE}
pred <- predict(knnFit, stdntmatTest, type="class")
tab <- table(true = stdntmatTest$Dalc, predicted = pred)
```


Calculating accuracy:
```{r, cache=TRUE}
sum(diag(tab)) / sum(tab)
```


#Optimal k and performance
Now we can find the opitmal value of `k` in the k-means algorithm.
For this task we can assume that the best value would be the one which yields the highest value for performance.
```{r, cache=TRUE}
tuneK <- 1:100
performance <- sapply(tuneK, function(k) {
  knnFit <- knn3(Dalc ~ studytime + age + absences + failures + freetime + goout, data = stdntmatTrain, k=k)
  tab <- table(true = stdntmatTest$Dalc,
          predict = predict(knnFit, stdntmatTest, type="class"))
  tab2 <- prop.table(tab, 1)
  tab2
  sum(diag(tab)) / sum(tab)
}) 
```


Optimal k and its performance:
```{r, cache=TRUE}
optimal_k = which.max(performance)
optimal_k
performance[optimal_k]
```


#Performance plot
```{r, warning=FALSE, cache=TRUE}
df <- data.frame(tuneK, performance)

ggplot(df, aes(tuneK, performance)) +
  geom_point() + 
  geom_smooth(se=FALSE, span=0.1, size=2) +
  theme_bw()

```
