---
title: "Homework #6 - LDA & QDA"
author: "Karolina Kwasiborska"
date: "2015-11-19"
output: 
  html_document:
    toc: TRUE
---

# The Homework

Use the k-fold cross validation to assess the performance of lda/qda on the wines dataset.

## Loading data
```{r}
library(caret)

wines <- read.table("http://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-red.csv", sep=";", header=TRUE)
wines$quality <- factor(ifelse(wines$quality > 5, "good", "bad")) 

train_control <- trainControl(method="cv", number=10)
```

## Linear discriminant analysis (LDA)
```{r}
lda.model<- train(quality~., method='lda',preProcess=c('scale', 'center'), data=wines, trControl=train_control)
lda.pred <- predict(lda.model, wines)

confusionMatrix <- confusionMatrix(wines$quality, lda.pred)$overall
round(confusionMatrix*100,2)
```

## Quadratic discriminant analysis (QDA)
```{r}
qda.model<- train(quality~.,   method='qda', preProcess=c('scale', 'center'), data=wines, trControl=train_control)
qda.pred <- predict(qda.model, wines)

confusionMatrix <- confusionMatrix(wines$quality, qda.pred)$overall
round(confusionMatrix*100,2)
```

## Summary
The performace of LDA and QDA on the wines dataset is not high.

LDA: Average accuracy - 74.55, Kappa - 48.98

QDA: Average accuracy - 74.42, Kappa - 48.21

LDA gives a little better result but difference is not significant.
