---
title: "Data exploration"
author: "Przemyslaw Biecek"
date: "Data Mining - Advances"
output: 
  html_document:
    toc: TRUE
---

# Introduction

In this part we are going to cover following topics:

* how to read data into R
* how to present basic summaries for vectors
* how to present summaries for groups

We are going to work with thee datasets: `GermanCredit` from `caret` package, `pancan` from csv/xlsx file, `iris` from `datasets` package.

# Loading data

## Data from core packages

For datasets that are available in base packages it's enough to use the `data()` function in order to load the data into the namespace.

Here we are going to load famous Anderson's iris data.

```{r}
data(iris)
head(iris)
```

## Data from additional packages

For datasets that are available in other packages load the package and then all datasets are available.

```{r}
library(caret)
data(GermanCredit)

head(GermanCredit[,1:10])
```

## Data from cvs/rda/xlsx files

Use `read.table{}` or `read.xlsx{openxlsx}` in order to read data from csv/xlsx files. 
Use `load{}` function in order to load data from binary rda files.

Here we are going to read data from The Cancer Genome Atlas. For 5 different cancers we have information about expression of 31 different genes.

```{r}
load("pancan.rda")
pancan <- read.table("pancan.csv", header=TRUE, sep=";")
library(openxlsx)
pancan <- read.xlsx("pancan.xlsx",sheet = 1)

head(pancan)
```

# Summaries for one variable

## Numerical summaries for one variable

Let's start with contingency tables for categorical variables like `disease` or `gender` from the `pancan` dataset.

The `table()` function calculates contingency tables for given variable.

```{r}
# species in the iris dataset
table(iris$Species)

# classes of credit Good/Bad
table(GermanCredit$Class)

# diseases/genders in pancan dataset
table(pancan$disease)
table(pancan$gender)
```

For continuous variable one can use the `summary()` function. 

```{r}
# 5 quartiles and mean for Sepal.Length
summary(iris$Sepal.Length)

# Age for German Credit data
summary(GermanCredit$Age)

# expression for selected genes
summary(pancan$ARHGAP8)
summary(pancan$PAX8)
```

## Graphical summaries for one variable

In many cases it is a good ide to present summaries graphically.

Contingency tables may be presented by barplots. 
Play with additional arguments like `horiz` or `las` to get nicer graph.

```{r}
# classes of credit Good/Bad
tab_credits <- table(GermanCredit$Class)
barplot(tab_credits)

# diseases/genders in pancan dataset
tab_disease <- table(pancan$disease)
par(mar=c(5,15,2,2))
barplot(tab_disease, horiz = TRUE, las=1)
```

For continuous variable distributions may be presented with histograms `hist()` or boxplots `boxplot()`.

```{r}
# Age for German Credit data
par(mar=c(5,5,2,2))
hist(GermanCredit$Age, col="grey")
boxplot(GermanCredit$Age, horizontal = TRUE)

# expression for selected genes
hist(pancan$ARHGAP8, col="grey")
boxplot(pancan$ARHGAP8, horizontal = TRUE)
```

# Summaries for two variables

## Numerical summaries for two variables

The `table()` function may be used for two categorical variables as well. 

Use `prop.table()` to calculate proportions in rows / columns.

```{r}
# classes of credit Good/Bad
tab <- table(GermanCredit$Class, GermanCredit$Personal.Male.Single)
tab
prop.table(tab, 1)

# diseases/genders in pancan dataset
tab2 <- table(pancan$disease, pancan$gender)
tab2
round(100*prop.table(tab2, 1),1)
```

Use `by()` function to calculate summaries for different sub-groups. 

```{r}
# Age for German Credit data
by(GermanCredit$Age, GermanCredit$Class, summary)

# expression for selected genes
by(pancan$ARHGAP8, pancan$disease, summary)
```

## Graphical summaries for two variables

Two dimensional contingency tables may be presented with mosaicplots.
They are not easy to read but very useful.

```{r}
tab <- table(GermanCredit$Class, GermanCredit$Personal.Male.Single)
tab
mosaicplot(tab)

# diseases/genders in pancan dataset
tab2 <- table(pancan$disease, pancan$gender)
tab2
mosaicplot(tab2, las=2, main="", color = c("red", "blue"))
```

For continuous variable boxplots (`boxplot()`) are very good for making comparisons. 
Try `vioplot{vioplot}` if you are not scared by densities.

```{r}
# Age for German Credit data
boxplot(GermanCredit$Age ~ GermanCredit$Class)

# expression for selected genes
par(mar=c(5,15,2,2))
boxplot(pancan$ARHGAP8 ~ pancan$disease, horizontal = TRUE, las=1)

library(vioplot)
par(mar=c(5,5,2,2))
vioplot(pancan$ARHGAP8[pancan$disease == "breast invasive carcinoma"],
        pancan$ARHGAP8[pancan$disease == "glioblastoma multiforme"],
        pancan$ARHGAP8[pancan$disease == "kidney clear cell carcinoma"],
        pancan$ARHGAP8[pancan$disease == "ovarian serous cystadenocarcinoma"])
```


# Home work

Use knitr to create a report with numerical and graphical summaries 

