---
title: "Home work"
author: "Karolina Kwasiborska"
date: "2015-10-08"
output: 
  html_document:
    toc: TRUE
---
# Exercise

Use `knitr` to create a report with numerical and graphical summaries for GermanCredit.

Summarize following variables: Duration, Amount, Class, Personal.Male.Single.


# Loading data

Loading data GermanCredit from package "caret"

```{r}
library(caret)
data(GermanCredit)
```

# Summaries for one variable

## Numerical summaries for one variable

### Contingency tables:

Discrete variables
```{r}
table(GermanCredit$Class)
table(GermanCredit$Personal.Male.Single)
```

Continuous variables (divided into 4 intervals)
```{r}
table(cut(GermanCredit$Duration, 4))
table(cut(GermanCredit$Amount, 4))
```

### Summary for continuous variables (5 quartiles and mean):
```{r}
summary(GermanCredit$Duration)
summary(GermanCredit$Amount)
```

## Graphical summaries for one variable

### Barplots:

Visualization of continuous and discrete variables divided into categories

```{r}
table_class <- table(GermanCredit$Class)
barplot(table_class)
```

Results: there are twice as much 'Good' credits as 'Bad' ones

```{r}
table_male <- table(GermanCredit$Personal.Male.Single)
barplot(table_male)
```

Results: there are a little bit more male single people

```{r}
table_amount <- table(cut(GermanCredit$Amount, 6))
par(mar=c(5,15,2,2))
barplot(table_amount, horiz = TRUE, las=1)
```

Results: most of credits are small

```{r}
table_duration <- table(cut(GermanCredit$Duration, 6))
par(mar=c(5,15,2,2))
barplot(table_duration, horiz = TRUE, las=1)
```

Results: most of credits are taken up to 26 years

### Histrograms:

Visualisation of continuous variables arranged in ranges

```{r}
hist(GermanCredit$Duration, 20, col="blue")
hist(GermanCredit$Amount, 100, col="grey")
```

Results: results are similar to the results of barplots but more precise

### Boxplots:

Visualization of quartiles, mean and outliers

```{r}
boxplot(GermanCredit$Amount, horizontal = TRUE)
```

Results: there are a lot of outliers

```{r}
boxplot(GermanCredit$Duration, horizontal = TRUE)
```

# Summaries for two variables

## Numerical summaries for two variables

### Proportion tables:

Proportions of each variable in terms of 'Good' and 'Bad' credits
```{r}
tab <- table(GermanCredit$Class, GermanCredit$Personal.Male.Single)
tab
round(100*prop.table(tab, 1),1)

tab <- table(GermanCredit$Class, cut(GermanCredit$Amount, 6))
tab
round(100*prop.table(tab, 1),1)

tab <- table(GermanCredit$Class, cut(GermanCredit$Duration, 6))
tab
round(100*prop.table(tab, 1),1)

```

### Summaries for different sub-groups('Good' and 'Bad' credits):

```{r}
by(GermanCredit$Amount, GermanCredit$Class, summary)

by(GermanCredit$Duration, GermanCredit$Class, summary)
```

## Graphical summaries for two variables

### Mosaicplots:

Visualization of relations between variables

```{r}
tab <- table(GermanCredit$Class, GermanCredit$Personal.Male.Single)
tab
mosaicplot(tab, las=2, main="Personal.Male.Single / Class", color = c("red", "blue"))

tab <- table(GermanCredit$Class, cut(GermanCredit$Amount, 4))
tab
mosaicplot(tab, las=2, main="Amount / Class", color = c("red", "blue"))

tab <- table(GermanCredit$Class, cut(GermanCredit$Duration, 4))
tab
mosaicplot(tab, las=2, main="Duration / Class", color = c("red", "blue"))

```

### Boxplots:

Comparisons of boxplots in terms of 'Good' and 'Bad' credits

```{r}
boxplot(GermanCredit$Amount ~ GermanCredit$Class, horizontal = TRUE, las=1)
boxplot(GermanCredit$Duration ~ GermanCredit$Class, horizontal = TRUE, las=1)
```

### Vioplot:

```{r}
library(vioplot)
par(mar=c(5,5,2,2))
vioplot(GermanCredit$Amount[GermanCredit$Class == "Good"],
        GermanCredit$Amount[GermanCredit$Class == "Bad"])
```