---
title: "P21"
author: "Klaudia Magda"
date: "9 grudnia 2016"
output: html_document
---

#Introduction
In given data set we will make an Extrapolatory Data Analysis.

We will answer following questions:

What are the values in given dataset? 
What are the distributions?
Are there relationships? Where? 

What kind of data types?(continuous/discreet/categorical)

Identify outliers
Goal of 1st part of project:graphical displays of data and numerical summaries of data



#Libraries
```{r, message=FALSE, warning=FALSE}
library(xlsx)
library(openxlsx)
library(psych)
library(e1071)
library(caret)
library(magrittr)
library(plyr)
library(dplyr)
library(party)
library(rpart)
```



#Loading data
```{r, message=FALSE, warning=FALSE, include=FALSE}

dataForm1 <- read.xlsx("C:/Users/kmagda001/Documents/Data-project2-CNK/dane_kwestionariuszowe.xlsx", 1, colNames=TRUE)

dataForm2 <- read.xlsx("C:/Users/kmagda001/Documents/Data-project2-CNK/dane_kwestionariuszowe.xlsx", 2, colNames=TRUE)

dataObserv1<- read.xlsx("C:/Users/kmagda001/Documents/Data-project2-CNK/dane_obserwacyjne.xlsx", sheet = 1, colNames = TRUE)

#dataObserv2<- read.xlsx("C:/Users/kmagda001/Documents/Data-project2-CNK/dane_obserwacyjne.xlsx", sheet=2, colNames = TRUE)


```


#Data Preprocessing

First step in data Preprocessing is to drop unnecessary columns that do not have huge impact on our data. 

Predictors that has been dropped:
`Kolumn1` and `lp` (no information about those predictors)
Columns that has majority of NA values (more than 85%)

In other columns values NA were replaced with certain value.


```{r, echo=TRUE, message=FALSE, warning=FALSE}

describe(dataObserv1)

dataObs <- subset(dataObserv1, select = -c(Kolumna1, lp, X3_OSTOW, X4_OSTOW, X5_OSTOW, X6_OSTOW, X7_OSTOW, X8_OSTOW, X9_OSTOW, uwagi))

describe(dataObs)
summary(dataObs)
ls.str(dataObs)

dataForm<-subset(dataForm1, select = -c(Kolum1))

dataMerge<-merge(dataObs,dataForm)

dataObs$opis[is.na(dataObs$opis)] <- 0
dataObs$animator[is.na(dataObs$animator)] <- 0

```




#Division on Categorical and Numerical predictors
Need to assign factor variable to categories, that were mentioned in description
```{r, echo=TRUE, message=FALSE, warning=FALSE}

colsObs <- c("ID","kategorie", "zach")
dataObs[,colsObs] <- data.frame(apply(dataObs[colsObs], 2, as.factor))


split(names(dataObs),sapply(dataObs, function(x) paste(class(x), collapse=" ")))


colsForm <- c("Plec", "studiaM", "studiaT", "pracaM" ,"pracaT","opis","animator", "NR_szkoły","NR_klasy","NR_dziennik")
dataForm[,c(5:12, 1)] <- data.frame(apply(dataForm[5:12], 2, as.factor))


split(names(dataForm),sapply(dataForm, function(x) paste(class(x), collapse=" ")))


colsMerge <- c("ID", "kategorie", "zach", "Plec","studiaM", "studiaT", "pracaM" ,"pracaT","opis","animator", "NR_szkoły","NR_klasy","NR_dziennik")
dataMerge[,colsMerge] <- data.frame(apply(dataMerge[colsMerge], 2, as.factor))


split(names(dataMerge),sapply(dataMerge, function(x) paste(class(x), collapse=" ")))


```

#Number of features

```{r, echo=TRUE, message=FALSE, warning=FALSE}

dataForm$oceJP[is.na(dataForm$oceJP)] <- 1
dataForm$oceM[is.na(dataForm$oceM)] <- 1
dataForm$oceP[is.na(dataForm$oceP)] <- 1


Eksponats<-length(unique(dataMerge$ekspot))
Eksponats
People<-length(unique(dataMerge$ID))
People

Top5People<-tail(names(sort(table(dataObs$ID))), 5)
Top5People

#aggregate(dataObs[, 1:2], list(dataObs$ekspot), mean)



#Average of Grades for each school
aggregate(dataForm[, 13:15], list(dataForm$NR_szkoły), mean)



aggregate(dataObs$czas_w_sek, list(dataObs$ekspot), mean)

```


#Summary Statistics
##Five number summaries (mean/median, min, max, q1, q3)
Five number summaries will show distribution of data and define if there are outliers variables
```{r, echo=TRUE, message=FALSE, warning=FALSE}

summary(dataObs$czas_w_min) # Summary statistics are printed for this variable


plot(density(dataObs$czas_w_sek), xlim=c(0,70), col="red", xlab="Time in seconds")


```

#Histogram
Histogram were used to notice if grades can be considered as outliner variables
```{r, echo=TRUE, message=FALSE, warning=FALSE}

attach(dataObs) 
attach(dataForm)

grades <- seq(0, 6, 1) 

hist(oceM, breaks=grades, xlab = "Grades", ylab = "Frequency", main = "Histogram of Math Grades", cex=0.1)

hist(oceJP, breaks=grades, xlab = "Grades", ylab = "Frequency", main = "Histogram of Polish Language Grades", cex=0.1)

hist(oceP, breaks=grades, xlab = "Grades", ylab = "Frequency", main = "Histogram of Biology Grades", cex=0.1)



```

#Boxplots

```{r, echo=TRUE, message=FALSE, warning=FALSE}
boxplot(oceM, bty="n",xlab = "Math Grades", cex=0.1)
boxplot(oceJP, bty="n",xlab = "Polish Grades", cex=0.1)
boxplot(oceP, bty="n",xlab = "Biology Grades", cex=0.1)
```






```

#Scatterplots
Scatterplots are used to visualize data.

Categories of exhibits:

1- breaks
2 - exhibits
3 - other atractions

Categories of behavior:

1 - Observation
2 - Touching
3 - Using
4 - Experiments

Categories of Gender:

1 - Girls
2 - Boys

Parents studies/work:
1 - Yes 
2 - No
3 - I dont know

```{r, echo=TRUE, message=FALSE, warning=FALSE}

 pairs(dataForm[,10:15], main = "Grades", pch = "+", col = c("red", "green3",  "blue"))


pairs(dataObs[,4:9], gap=0, pch=19, cex=0.4, col=c("darkblue","red"))




xyplot(czas_w_sek ~ start_s | kategorie, dataObs, groups = dataObs$kategorie, pch= 20)
xyplot(start_s ~ stop_s | kategorie, dataObs, groups = dataObs$kategorie, pch= 20)
xyplot(start_s ~ stop_s | zach, dataObs, groups = dataObs$zach, pch= 20)

xyplot(start_s ~ czas_w_sek | zach, dataObs, groups = dataObs$zach, pch= 20)



dataMerge$Plec<-as.factor(dataMerge$Plec)
dataMerge$kategorie<-as.factor(dataMerge$kategorie)
dataMerge$oceP<-as.factor(dataMerge$oceP)
dataForm$Plec<-as.factor(dataForm$Plec)
#dataForm$OceM<-as.factor(dataForm$OceM)
dataForm$oceJP<-as.factor(dataForm$oceJP)
#dataForm$OceP<-as.factor(dataForm$OceP)

spineplot(dataMerge$Plec,dataMerge$kategorie, col="green")

spineplot(dataForm$Plec ~ dataForm$oceJP , col="lightblue", main="Polish Grades for Gender")
spineplot(dataForm$Plec ~ dataForm$oceM , col="red", main="Math Grades for Gender")
spineplot(dataForm$Plec ~ dataForm$oceP , col="lightgreen", main="Biology Grades for Gender")

spineplot(dataForm$studiaT ~ dataForm$pracaT , col="pink", main="Father studies/work")

spineplot(dataForm$studiaM ~ dataForm$pracaM , col="yellow", main="Mother studies/work")



#xyplot(start_s ~ czas_w_sek | Plec, data, groups = dataMerge$Plec, pch= 20)

#xyplot(start_s ~ stop_s | Plec, data, groups = dataMerge$Plec, pch= 20)
```
#XY plot


```{r, echo=TRUE, message=FALSE, warning=FALSE}

xyplot(czas_w_sek ~ start_s | kategorie, dataObs, groups = dataObs$kategorie, pch= 20)
xyplot(start_s ~ stop_s | kategorie, dataObs, groups = dataObs$kategorie, pch= 20)
xyplot(start_s ~ stop_s | zach, dataObs, groups = dataObs$zach, pch= 20)

xyplot(start_s ~ czas_w_sek | zach, dataObs, groups = dataObs$zach, pch= 20)
```
