---
title: "P22"
author: "Klaudia Magda"
date: "9 grudnia 2016"
output: html_document
---

#Introduction
In 2nd part of the project I will focus on a division of data into groups of similar objects that are similar between themselves and dissimilar to objects of other groups (Clustering). 

#Libraries
```{r, message=FALSE, warning=FALSE}
library(xlsx)
library(openxlsx)
library(psych)
library(e1071)
library(caret)
library(magrittr)
library(plyr)
library(dplyr)
library(party)
library(rpart)
library(psych)
library(ggplot2)
library(NbClust)
library(pvclust)
library(MASS)
library(bclust)
```



#Loading data
```{r, message=FALSE, warning=FALSE, include=FALSE}
dataForm1 <- read.xlsx("C:/Users/kmagda001/Documents/Data-project2-CNK/dane_kwestionariuszowe.xlsx", 1, colNames=TRUE)

dataForm2 <- read.xlsx("C:/Users/kmagda001/Documents/Data-project2-CNK/dane_kwestionariuszowe.xlsx", 2, colNames=TRUE)

dataObserv1<- read.xlsx("C:/Users/kmagda001/Documents/Data-project2-CNK/dane_obserwacyjne.xlsx", sheet = 1, colNames = TRUE)
```


#Data Preprocessing

First step in data Preprocessing is to drop unnecessary columns that do not have huge impact on our data. 

Predictors that has been dropped:
`Kolumn1` and `lp` (no information about those predictors)
Columns that has majority of NA values (more than 85%)
```{r, echo=TRUE, message=FALSE, warning=FALSE}
describe(dataObserv1)

dataObs <- subset(dataObserv1, select = -c(Kolumna1, lp, X3_OSTOW, X4_OSTOW, X5_OSTOW, X6_OSTOW, X7_OSTOW, X8_OSTOW, X9_OSTOW, uwagi))

describe(dataObs)
summary(dataObs)
ls.str(dataObs)

dataForm<-subset(dataForm1, select = -c(Kolum1, wyk_zawod_M, wyk_zawod_T, aspiracje))

colsObs <- c("ID","kategorie", "zach")
dataObs[,colsObs] <- data.frame(apply(dataObs[colsObs], 2, as.factor))

colsForm <- c("Plec", "studiaM", "studiaT", "pracaM" ,"pracaT","opis","animator", "NR_szkoły","NR_klasy","NR_dziennik")
dataForm[,c(5:12)] <- data.frame(apply(dataForm[5:12], 2, as.factor))


split(names(dataForm),sapply(dataForm, function(x) paste(class(x), collapse=" ")))
```

#NA values
To deal with NA values I have decided to drop those variables, where it is hard to predict NA value and in other cases assign most frequent value for categorical variables and average for numerical.
```{r, echo=TRUE, message=FALSE, warning=FALSE}
dataObs <- subset(dataObs, select = -c(Z_KIM, X1_OSTOW, X2_OSTOW))


for(i in 1:ncol(dataObs)){
  dataObs[is.na(dataObs$start_min), i] <- mean(dataObs[,i], na.rm = TRUE)
}

for(i in 1:ncol(dataObs)){
  dataObs[is.na(dataObs$start_s), i] <- mean(dataObs[,i], na.rm = TRUE)
}

for(i in 1:ncol(dataObs)){
  dataObs[is.na(dataObs$stop_min), i] <- mean(dataObs[,i], na.rm = TRUE)
}

for(i in 1:ncol(dataObs)){
  dataObs[is.na(dataObs$stop_s), i] <- mean(dataObs[,i], na.rm = TRUE)
}


dataObs$zach[is.na(dataObs$zach)]
dataObs$opis[is.na(dataObs$opis)] <- 0
dataObs$animator[is.na(dataObs$animator)] <- 0
dataObs$kategorie[is.na(dataObs$kategorie)] <- 2
dataObs$zach[is.na(dataObs$zach)] <- 3

dataForm$studiaM[is.na(dataForm$studiaM)] <- 3
dataForm$studiaT[is.na(dataForm$studiaT)] <- 3
dataForm$pracaM[is.na(dataForm$pracaM)] <- 3
dataForm$pracaT[is.na(dataForm$pracaT)] <- 3
dataForm$oceM[is.na(dataForm$oceM)] <- 1
dataForm$oceJP[is.na(dataForm$oceJP)] <- 1
dataForm$oceP[is.na(dataForm$oceP)] <- 1

colnames(dataObs)[colSums(is.na(dataObs)) > 0]
colnames(dataForm)[colSums(is.na(dataForm)) > 0]


dataForm<-na.omit(dataForm)
dataObs<-na.omit(dataObs)


#split(names(dataMerge),sapply(dataMerge, function(x) paste(class(x), collapse=" ")))
```


#Kmeans
Kmeans has been implemented for 3 and 5 clusters. Moreover, I have decided to make a clustering for time in seconds (start time, stop time and duration) and from grades in dataForm.
```{r, echo=TRUE, message=FALSE, warning=FALSE}
library(cluster)

time <- c("start_s","stop_s", "czas_w_sek")
grades <- c("oceM","oceJP", "oceP")


clusGap(dataObs[,time], kmeans, 10, B = 100, verbose = interactive())
clusGap(dataForm[,grades], kmeans, 10, B = 100, verbose = interactive())


means <- kmeans(dataObs[,time], 3)
dataObs$cluster <- factor(kmeans$cluster)

ggplot(dataObs, aes(start_s, stop_s, czas_w_sek)) +
  geom_text(size=3, aes(label=ID, color=cluster)) + 
  geom_point(data=dataObs, size=3)+
  theme_bw()


means1 <- kmeans(dataObs[,time], 5)
dataObs$cluster <- factor(kmeans$cluster)

ggplot(dataObs, aes(start_s, stop_s, czas_w_sek)) +
  geom_text(size=3, aes(label=ID, color=cluster)) + 
  geom_point(data=dataObs, size=5)+
  theme_bw()



means2 <- kmeans(dataForm[,grades], 3)
dataForm$cluster <- factor(kmeans$cluster)

ggplot(dataForm, aes(start_s, stop_s, czas_w_sek)) +
  geom_text(size=3, aes(label=ID, color=cluster)) + 
  geom_point(data=dataForm, size=3)+
  theme_bw()



means3 <- kmeans(dataForm[,grades], 5)
dataForm$cluster <- factor(kmeans$cluster)

ggplot(dataForm, aes(start_s, stop_s, czas_w_sek)) +
  geom_text(size=3, aes(label=ID, color=cluster)) + 
  geom_point(data=dataForm, size=5)+
  theme_bw()


```



#Distance partitioning
Distance 
```{r, echo=TRUE, message=FALSE, warning=FALSE}

d1<-dist(dataForm[,grades], method="euclidean")
fit1<-hclust(d1, method="ward")
plot(fit1)
groups <-cutree(fit2, k=3)
rect.hclust(fit2, k=3, border="red")


d2<-dist(dataForm[,grades], method="euclidean")
fit2<-hclust(d2, method="ward")
plot(fit2)
groups <-cutree(fit2, k=5)
rect.hclust(fit2, k=5, border="blue")


d3<-dist(dataObs[,time], method="euclidean")
fit1<-hclust(d3, method="ward")
plot(fit3)
groups <-cutree(fit3, k=3)
rect.hclust(fit3, k=3, border="green")


d4<-dist(dataForm[,time], method="euclidean")
fit4<-hclust(d4, method="ward")
plot(fit4)
groups <-cutree(fit4, k=5)
rect.hclust(fit4, k=5, border="orange")


```

