---
title: "Data_Mining_PROJECT"
author: "Shilpa Khichar & Elham Shahrour"
date: "9 January 2017"
output: html_document
---

The Dataset is a transaction set, it contains the items that are sets of items belonging to same itemset.
A-Priori algorithm is a level wise itemset mining technique. 

Fetching data
```{r read data}
#rm(list=ls())
table <- read.csv("F:/B.I.Sem2/data_mining/Project/Transaction_data.csv", header=FALSE)
tbl <- as.matrix(table)
colnames(tbl) <- NULL
itemsets <- tbl
items <- c(1,2,3,4,5)
min_sup <- 0.20*nrow(itemsets) 
```
We have 5 item in our dataset namely 1,2,3,4,5...we have to check which item are bought in pairs and in what order.
Defining minimum support as 20% of records in dataset/itemset

A-Priori Algorithm, have 2 subproblems:
1) Generation Of frequent ItemSets
2) Rule Generation

The Input item set is binarized, i.e. 1 means item is present in ItemSet else is absent for value =0


#	Algorithm: Apriori 
		generate the frequent itemsets from the input datasets with the specified support
	Input:
		data,as a MATRIX
		base_items,as VECTOR
		MIN_SUP, FLOATING point value
		parameter,
		
```{r apriorifunction}
Apriori <- function(data,base_items,MIN_SUP,parameter=NULL){
#  fetching Intial Candidate set 
	f <- InitCandidateSet(data,base_items)
# creating c as an empty list	
	c <- list()
# assigning fequent item set to the list c using FindFrequentItemset function	
	c[[1]] <- FindFrequentItemset(f,base_items,1,MIN_SUP)
	k <- 2
#	len4data <- GetDatasetSize(data)
	while( !IsEmpty(c,k-1) ){
		f[[k]] <- AprioriGen(c,k-1) 
		if(length(f)==k){
# incrementing at each iteration the count of support for each frequent item sets in the list c		  
			f[[k]] <- IncreaseSupportCount(f[[k]],data) 
			c[[k]] <- FindFrequentItemset(f,base_items,k,MIN_SUP)		
		}else{break}
		k <- k+1
	}
	c
}
```
next we define function AprioriGen to get the maximum frequent itemset(MFI)
F is the Maximum frequent itemset.
```{r apriprigenfunc}
AprioriGen <- function(c,k){
	ck <- c[[k]][,-ncol(c[[k]])]
	f <- NULL
	len <- nrow(ck)
	for(idx in seq(nrow(ck))){
		jdx <- idx+1
		while(idx<jdx && jdx<=len){
			a <- ck[idx,]
			b <- ck[jdx,]
			if( k==1 || identical(a[1:(k-1)],b[1:(k-1)]) ){
					ab <- ifelse(a+b,1,0)
# calling needPrune function to remove non frequent item set from list ck	and if the item doesnot require pruning then appending it to f.				
					if( !NeedPrune(ck,ab,k) ){
						f <- rbind(f,ab)
					}else{
						#print("Pruned")
					}
			}
			jdx <- jdx + 1
		}
	}

	if(length(f)){
		 f <- cbind(f,rep(0,dim(f)[1]))
		 rownames(f) <- NULL
	}
	return(f)
}
```
```{r prunefunc}
#checking if some itemset combination is not frequent then we remove/prune that item from our frequent item set list
NeedPrune <- function(ck,ab,k){
	ck <- rbind(ck,ab)
	len <- dim(ck)[1]
	for(idx in which(ab>0)){
		temp <- ab
		temp[idx] <- 0
		for(idx in seq(len)){
			if(identical(temp,ck[idx,]))break
		}
		if(idx==len)return(TRUE)
	}
	return(FALSE)
}
## increasing support count for the item sets that are in the frequency table (fk) calculatted previously
IncreaseSupportCount <- function(fk,data){
	w4f <- ncol(fk)
	len4f <- nrow(fk)
	len4d <- nrow(data)
	for(idx in seq(len4d)){
		for(jdx in seq(len4f)){
			if(identical(fk[jdx,-w4f],fk[jdx,-w4f]*data[idx,])){
				fk[jdx,w4f] <- fk[jdx,w4f] + 1
			}
		}
	}
	return(fk)
}

IsEmpty <- function(ck,k){
	return(ifelse(nrow(ck[[k]])>0,FALSE,TRUE))
}

GetDatasetSize <- function(data){
	return( nrow(data) )
}

InitCandidateSet <- function(data,base_items){
	list(cbind(diag(length(base_items)),apply(data,2,sum)))
}

FindFrequentItemset <- function(fk,base_items,k,MIN_SUP){
	data <- fk[[k]]
	return(data[data[,dim(data)[2]]>MIN_SUP,])
}
```
```{r func call}
frequent_itemsets <- Apriori(itemsets,items,min_sup)
print(frequent_itemsets)
```

LHS function is diving our datset into 2 parts and giving us the items that are bought by Customers after purchase of item number 2
```{r apriori rules}
matrix_tbl <- as.matrix(table)
LHS <- function(lhs, tbl,ans_list){
matrix_tbl <- as.matrix(tbl)
new_matrix <- matrix_tbl[matrix_tbl[,lhs] == 1,]
row <- nrow(new_matrix)
col <- ncol(new_matrix)
x <- list()
for(r in 1:row){
  itemset = c()
  i = 1
  for (c in 1:col){
    if(new_matrix[r,c] == 1) {
      itemset[i] = c
      i = i+1
      }
  }
  x[[r]] <- itemset
}
   ans_list =list()
  idx =1
  for(i in 1:length(x)){
     vector_x <- (x[[i]])
    position <- which(vector_x == lhs)
    len <- length(vector_x)
    len
    if(len > position){
     ans <- vector_x[c(position:len)]
     ans_list[[idx]] = ans
     idx=idx+1
     }
  }
   return(unique(ans_list))
}
```

RHS function is diving our datset into 2 parts and giving us the items that are bought by Customers before purchase of item number 2 or what products lead to purchase of item 2(here)
```{r  rhs }
RHS <- function(rhs, tbl,ans_list){
matrix_tbl <- as.matrix(tbl)
new_matrix <- matrix_tbl[matrix_tbl[,rhs] == 1,]
row <- nrow(new_matrix)
col <- ncol(new_matrix)
x <- list()

for(r in 1:row){
  itemset = c()
  i = 1
  for (c in 1:col){
    if(new_matrix[r,c] == 1) {
      itemset[i] = c
      i = i+1
      }
  }
  
  x[[r]] <- itemset
}
  ans_list =list()
  idx=1
    for(i in 1:length(x)){
    vector_x <- (x[[i]])
    position <- which(vector_x == rhs)
    len <- length(vector_x)
    len
    if(position > 1){
     ans <- vector_x[c(1:position)]
      ans_list[[idx]] = ans
       idx=idx+1
     }
    }
  return(unique(ans_list))
}

```
```{r results_of_lhs_rhs}
ANSWER_RHS <- RHS(rhs =2 , tbl = table, ans_list)
ANSWER_RHS
ANSWER_LHS <- LHS(lhs =2 , tbl = table, ans_list)
ANSWER_LHS
```
In the PROB function we calculate and display the itemsets and there respective frequencies  that the shopkeeper wants to analyse 
```{r final_output}
prob <- function(answer,tbl ){
  
if(length(answer)>0) { 
              for(i in 1:length(answer)){
                    vector_x <- (answer[[i]])
                    len <- length(vector_x)
                    new_matrix <- matrix_tbl
                    for(idx in 1:len){
                                      item <- vector_x[c(idx)]
                                      new_matrix <- new_matrix[new_matrix[,item] == 1,]
                                    }
                              freq = nrow(new_matrix)
                              freq
                              if( is.null(freq)){
                                  a <- as.data.frame(new_matrix)
                                  a <- t(a)
                                  freq <- nrow(a)}
                              print(c("list ->", vector_x , "frequency->", freq))
                                   }
}
else {print("No frequencies to show  in this case") }  
}

final_output <- prob(ANSWER_LHS, table)

```
```{r result}
table
ANSWER_RHS <- RHS(rhs =3 , tbl = table, ans_list)
ANSWER_RHS
ANSWER_LHS <- LHS(lhs =3 , tbl = table, ans_list)
ANSWER_LHS
final_output <- prob(ANSWER_LHS, table)
final_output <- prob(ANSWER_RHS, table)

ANSWER_RHS <- RHS(rhs =4 , tbl = table, ans_list)
ANSWER_RHS
ANSWER_LHS <- LHS(lhs =4 , tbl = table, ans_list)
ANSWER_LHS
final_output <- prob(ANSWER_LHS, table)
final_output <- prob(ANSWER_RHS, table)

ANSWER_RHS <- RHS(rhs =1 , tbl = table, ans_list)
ANSWER_RHS
ANSWER_LHS <- LHS(lhs =1 , tbl = table, ans_list)
ANSWER_LHS
final_output <- prob(ANSWER_LHS, table)
final_output <- prob(ANSWER_RHS, table)

ANSWER_RHS <- RHS(rhs =5 , tbl = table, ans_list)
ANSWER_RHS
ANSWER_LHS <- LHS(lhs =5 , tbl = table, ans_list)
ANSWER_LHS
final_output <- prob(ANSWER_LHS, table)
final_output <- prob(ANSWER_RHS, table)


```


