---
title: "Homework_7"
author: "Ziad Al Bkhetan"
date: "November 27, 2015"
output: 
  html_document:
    toc : TRUE
---


# Data Loading

```{r}
library(caret)
trips <- read.table("D:\\MSc Computer science and informatics\\3rd semester\\Data Mining - Advances\\Kogle Website\\train.csv", sep=",", head=T)
head(trips)

```

# Data Preparation
I will use dplyr library to prepare the dataset, byy grouping the records using the visitnumber, and each record will contain :
<br>the number of baught products from each department
<br>the count of returned products
<br>the count of baught products
<br>different finelNumber
<br>different departments count
 
```{r}
library(tidyr)
library(dplyr)
                                    
newTrips = trips %>% 
        group_by(VisitNumber) %>%
        summarise(
          TripType = head(TripType, 1),
          day = unique(Weekday)[1], 
          ProductsCount = sum(ScanCount >=0),
          unique_FinelineNumber = length(unique(FinelineNumber)),
          no_of_returned_products = sum(abs(ScanCount < 0)),
          unique_departments = length(unique(DepartmentDescription))
        )                            


departments = trips %>% 
  group_by(VisitNumber,TripType, DepartmentDescription) %>% 
  summarise(count = n()) %>%
  spread(DepartmentDescription, count, fill=0)

finalTrips <- merge(newTrips, departments, by=c("TripType", "VisitNumber"))
finalTrips = finalTrips[1:30000,]
                                    
finalTrips$TripType = factor(finalTrips$TripType)
names(finalTrips) = make.names(names(finalTrips), unique = FALSE)
                                    
library(caret)
indxTrain <- createDataPartition(y = finalTrips$TripType, p = 0.75)
finalTripsTrain <- finalTrips[indxTrain$Resample1,]
finalTripsTest <- finalTrips[-indxTrain$Resample1,]
                                                                     
```

# Apply Random Forests Classifier
I will apply random forests to classify this dataset, and calculate the error for this classifier
```{r}
library(randomForest)
rfTree<- randomForest(TripType~., data=finalTripsTrain)
confusionMatrix = confusionMatrix(finalTripsTest$TripType, predict(rfTree, finalTripsTest))$overall
round(confusionMatrix*100,2)
scores <- predict(rfTree, finalTripsTest, type = "prob")
myScores <- sapply(1:nrow(finalTripsTest), function(i){
  scores[i, finalTripsTest$TripType[i]]
})
mean(-log(pmax(myScores,0.05)))
                            
```
#Remarks
after apply this classifier different times I got different results approximatly about 0.56, but each time I increase the size of the dataset t get worse result

