---
title: "Homework 9"
author: "Ziad Al Bkhetan"
date: "December 10, 2015"
output: 
  html_document:
    toc : TRUE
---

#Introduction
In this Homework I will cluster the deputies from the biggest two parties depending on the voting, using two different distance calculations, and two different linkage methods, then I will visualize the data. 

# Data Loading
```{r , cache=FALSE, warning=FALSE, message=FALSE}
library(dplyr)
library(tidyr)
library(cluster)
library(ape)
library(RColorBrewer)

load("D:/MSc Computer science and informatics/3rd semester/Data Mining - Advances/Homeworks/all_votes.rda")

```

# Data Preparation
firstly I will filter the data to keep the records for the biggest two parties "PO and PiS".
<br>Then I will count the distinct deputies to filter the votes which not exceed 75 % of the deputies.
<br> to count the votes I will group by id_voting, then I will filter the data to get the final dataset I will use for clustering.
 
```{r , cache=FALSE, warning=FALSE, message=FALSE}
# filtering based on club value
mydata = all_votes[all_votes$club %in% c('PO', 'PiS'),1:7]

# get unique deputies according to the name and the club, because maybe the same name in different club
unique_deputies <- unique(mydata[c("surname_name", "club")])

# group votes

voting_count = mydata %>% 
  group_by(id_voting) %>%
  summarise(
    vote_count = n()
  )     

# find used votes in clustering
filtered_vote = voting_count[voting_count$vote_count > 0.75 * length(unique_deputies),]
# filter the dataset using the important votes "after filtration" 
mydata <- mydata[mydata$id_voting %in% filtered_vote$id_voting,]

# vote type coding
all_votes_new = mydata %>%
  mutate(
    voting_type = (ifelse(vote=='Against', -1, ifelse(vote=='For', 0, 1 )))
  )

all_votes_new = all_votes_new[,c("surname_name","club", "id_voting","voting_type")]

# create the final dataset using spread function
final_data = all_votes_new %>%
  spread(id_voting, voting_type, fill=0)


```

# Distance Matrix Calculation
I will use two different functions to calculate the distance using Id_voting values as coordinates:
<br> 1- euclidian distances
<br> 2- Manhattan distances

```{r , cache=FALSE, warning=FALSE, message=FALSE}
distinct_id_voting = unique(all_votes$id_voting)
rownames(final_data) = paste(final_data$club, final_data$surname_name, sep="_")

# euclidian distances for original dataset
mat1 <- dist(final_data[,distinct_id_voting])
final_mat = as.matrix(mat1)


# Manhattan distances for original dataset
mat2 <- dist(final_data[,distinct_id_voting], method = 'manhattan')
final_mat2 = as.matrix(mat2)

```

# Clustering
I will cluster the data using two different linkage method
<br> 1- average, I will apply it on the distance matrix I created using the euclidian distance function
<br> 2- complete, I will apply it on the distance matrix I created using the manhattan distance function
```{r , cache=FALSE, warning=FALSE, message=FALSE}
final_data$colr <- factor(ifelse(final_data$club == 'PO', 1, 2)) 
final_data2 <-final_data
# Using Average
hc <- agnes(final_mat, method="average")
plot(hc, which.plots=2, cex=0.5, main="")
final_data$labels = factor(cutree(hc, k=4))

# Using complete
hc2 <- agnes(final_mat2, method="complete")
plot(hc2, which.plots=2, cex=0.5, main="")
final_data2$labels = factor(cutree(hc2, k=4))
```

# Clusters Visualization 
I have clustered all the deputies and used their names and club as the row name.
<br> Red Color for PO party
<br> Blue Color for PiS party

## Clusters Using Euclidian Distance And Average Linkage
```{r , cache=FALSE, warning=FALSE, message=FALSE}

cols <- brewer.pal(3,"Set1")
hc <- as.phylo(as.hclust(agnes(final_data, method="complete")))

par(mar=c(1,1,2,1), xpd=NA)

plot(hc, type = "fan", cex = 0.8,
     tip.color = cols[final_data$colr])

plot(as.phylo(hc), type = "unrooted", cex = 0.8,
     tip.color = cols[final_data$colr])

plot(as.phylo(hc), type = "radial", cex = 0.8,
     tip.color = cols[final_data$colr])

plot(as.phylo(hc), type = "phylogram", cex = 0.8,
     tip.color = cols[final_data$colr])

plot(as.phylo(hc), type = "cladogram", cex = 0.8,
     tip.color = cols[final_data$colr])


```

## Clusters Using Manhattan Distance And Complete Linkage
```{r , cache=FALSE, warning=FALSE, message=FALSE}
cols <- brewer.pal(3,"Set1")
hc2 <- as.phylo(as.hclust(agnes(final_data2, method="complete")))

par(mar=c(1,1,2,1), xpd=NA)

plot(hc2, type = "fan", cex = 0.8,
     tip.color = cols[final_data2$colr])

plot(as.phylo(hc2), type = "unrooted", cex = 0.8,
     tip.color = cols[final_data2$colr])

plot(as.phylo(hc2), type = "radial", cex = 0.8,
     tip.color = cols[final_data2$colr])

plot(as.phylo(hc2), type = "phylogram", cex = 0.8,
     tip.color = cols[final_data2$colr])

plot(as.phylo(hc2), type = "cladogram", cex = 0.8,
     tip.color = cols[final_data2$colr])


```

# Remarks
Because of the big number of deputies, the visualization will not be clear, and we can filter the deputies more to get more clear results.