---
title: "Internet of Things"
author: "Anand Gowrishankar Sudhan "
date: "8 January 2016"
output: html_document
---

```{r, echo=FALSE}
#load("SmallLogs.rda") how you should be loading the data 
#in my case I should load it in csv.
SmallLogs <- read.table("smallLogs_170000.csv",header=T,sep=',')
```

#Phase I : Exploratory Data Analysis(EDA)

##Putting the data in the correct form

```{r, message=FALSE,warning=FALSE}
#Elliminate the first var(useless indexing and memory consuming)
SmallLogs <- SmallLogs[,-1]

# Setting date as POSIXct object that we can do compuations over it 
SmallLogs$time <- as.POSIXct(SmallLogs$date,format="%Y-%m-%d %H:%M:%S")

#Extract scense names from the variable "scene" and convert them as factors 
test=sub(".*=", "", SmallLogs$scene)
SmallLogs$scene_name<- substr(test,1, nchar(test)-1)
SmallLogs$scene_name<-as.factor(SmallLogs$scene_name)
```


#Calculating time of interaction 
```{r, message=FALSE,warning=FALSE}
library(dplyr)
#first log to the station
test=SmallLogs %>% group_by(station,visitor) %>% summarise(first=first(time)) 
 #last log to the station
test2=SmallLogs %>% group_by(station,visitor) %>% summarise(last=last(time)) 
dataset=merge(test,test2,by=c("station","visitor"))
#Time of interaction
dataset$time_inter=as.numeric(dataset$last-dataset$first)
```


###Distribtion of time of interaction in seconds

```{r, message=FALSE,warning=FALSE}
library(ggplot2)
ggplot(data=dataset,aes(x=time_inter))+geom_bar(binwidth=1)+xlim(0,1000)
summary(dataset$time_inter)
```

The time of interaction have the great concentration before 500s(8 min) and most of interactions lasts two min if outliers are eliminated.
Only 515 visitors have a time of interaction greater than 500s and it because of the structure of data(a lot of errors like entering two time successively without logging out)
Clearly, by showing numeric summaries of the time of interaction, the max value and the third quartile are way too separated.

We aggregate stations according to their names, for example stations **chnk01 ... chnk09** will be grouped into one group *chnk0* and so on.

```{r}
#Agreggate stations for better visualisations 
index_0=grep("cnk0",as.character(dataset$station))
index_1=grep("cnk1",as.character(dataset$station))
index_2=grep("cnk2",as.character(dataset$station))
index_3=grep("cnk3",as.character(dataset$station))
index_4=grep("cnk4",as.character(dataset$station))
index_5=grep("cnk5",as.character(dataset$station))
index_6=grep("cnk6",as.character(dataset$station))
index_7=grep("cnk7",as.character(dataset$station))

C=c(rep("cnk0",length(index_0)),rep("cnk1",length(index_1)),rep("cnk2",length(index_2)),rep("cnk3",length(index_3)),rep("cnk4",length(index_4)),rep("cnk5",length(index_5)),rep("cnk6",length(index_6)),rep("cnk7",length(index_7)))

dataset$stationAgg=C
```

###Distribution of time interaction by station

```{r, message=FALSE,warning=FALSE,fig.height=7,fig.width=14}
ggplot(data=dataset,aes(x=time_inter))+geom_bar(aes(y =(..count..)/sum(..count..)))+facet_grid(~station)+xlim(0,500)+xlab("Time of interaction(s)")+ylab("Percentage")
```

Time of interaction according to stations tend to have different distributions, for example stations **cnk02a** and **cnk02b** have a great interaction for the time 60 seconds.

```{r,message=FALSE,warning=FALSE}
#Boxplot of time interaction according to stations
ggplot(dataset,aes(y=time_inter,x=station))+geom_boxplot()+ylim(0,1000)
```

Using box plots we show the mean, *1st* quartile and *3rd* quartile for time interaction in each station.
Station  **cnk03** have a mean greater than other stations

##Relation between time of interaction, weekday and hour of the day

```{r, message=FALSE,warning=FALSE}
dataset$Dayofweek=as.factor(weekdays(dataset$first))
ggplot(dataset,aes(x=Dayofweek,y=time_inter/60))+geom_boxplot(alpha=1/3)+ylim(0,10)+ylab("Time of interaction(min)")
```

In the over all shape of distributions according to the time of interaction, we remark that days don't have a real influence on the time of interaction, unless the special case of Monday.
We remark also a great concentration of interaction whose time is less than *2.5min*.
Monday don't figure on the plot because of the fact that the machines aren't used heavily that day of the week(recall that we've used a subset, for the whole data set a very few logs are made on Monday).   
we can look at the table of count of days that visitors use the machines:

```{r, message=FALSE,warning=FALSE}
table(dataset$Dayofweek)
dataset=subset(dataset,Dayofweek!="Monday")
```
####  Using one way ANOVA to test the effect of day of the week over time of interaction
Let's compare means of time of interaction in each day of the week using anova.
```{r, message=FALSE,warning=FALSE}
aov=aov(time_inter ~ Dayofweek, data=dataset)
summary(aov)
```
The output from **the one way ANOVA table** confirm that there is an influence of the days of the week over time of interaction (pvalue < 0.05), and deny then our graphical findings. 

```{r, message=FALSE,warning=FALSE}
library(gplots)
attach(dataset)
plotmeans(time_inter ~ Dayofweek,xlab="Days of the week",
  ylab="Time of interaction in seconds", main="Mean plot of time interaction\nwith 95% CI")
detach(dataset)
```



```{r, message=FALSE,warning=FALSE}
#hours of the day 
dataset$hour<- format(dataset$first,"%H")

#Percentage for each hour of using machines
round(table(dataset$hour)*100/nrow(dataset),digits = 2)
```

Hour of logging does have influence on time interaction with stations, The period between 11 a.m and 4p.m is the part of the day that visitors interact with stations the most. 



**What features are used the most?**  
The next table will show the most used actions by visitors 
```{r, message=FALSE,warning=FALSE}
head(sort(table(SmallLogs$scene_name),decreasing = TRUE),10)
tb=(table(SmallLogs$scene_name))
tb=data.frame(tb)
sb=subset(tb,Freq>1500)
index=as.vector(sb[,1])
```

```{r, message=FALSE,warning=FALSE,fig.height=7,fig.width=14}
#This vizualisation is about the most features used
ggplot(data = sb,aes(x=Var1,y=Freq))+geom_bar(stat="identity")+theme(axis.text.x=element_text(angle=90,size=9))+xlab("Scene")+ylab("Count")
```
The actions (scenes or features): Splash, Main, RemovalWarningScence,Dance,Example are the most used actions by the users when they interact with machines.


# Phase II : Clustering visitors 

##Define similarity dissimilarity between visitors 

The first step is to find a way to characterize the dissimilarity between visitors. 
The relations that we've made to find links with descriptive statistics will help to define a profiling of users according to hour of logs, stations and patterns. To do so, and since
the variables we've got are for the most categorical, the data set is also huge to process v
via k modes (hard to be scalable) we decided to convert factors to dummy variables and then perform the k-means by scaling the resulting variables.

```{r, message=FALSE,warning=FALSE}
library(dplyr)

#Adding week of the day when the log took place and the hour of logging
SmallLogs$Dayofweek=as.factor(weekdays(SmallLogs$time))
SmallLogs$date=NULL
SmallLogs$hour <- format(SmallLogs$time,"%H")

Logstocluster=merge(SmallLogs,dataset[,c(1,2,5)],by=c("station","visitor"),all=TRUE)

#This selection ignore the day of the week
Logstocluster <- select(Logstocluster,-scene,-time,-Dayofweek)
Logstocluster$hour<-as.factor(Logstocluster$hour)

#Converting factors to dummy variables 
library(ade4)
test2=acm.disjonctif(Logstocluster[,-c(2,6)])
test1=select(Logstocluster,visitor,time_inter)

#final dataset
test2$visitor=test1$visitor
test2$time_inter=test1$time_inter
Logstocluster=test2
```

##Within Sum of Squared to choose **k**:

```{r, message=FALSE,warning=FALSE}
n=ncol(Logstocluster)
wss=data.frame(NULL)


#Scale variables
Logstocluster.s=scale(Logstocluster[,-c(n,n-1)])
set.seed(1234)
for (i in 2:15){
  wss[i,1]=sum(fit=kmeans(Logstocluster.s,centers=i)$withinss)
}
colnames(wss)="wss"
wss$k=seq(1,15,1)
ggplot(data = wss,aes(x=k,y=wss))+geom_line()+scale_x_discrete(breaks=seq(1,15,1))

```

According to the wss it seems that there is a great dissimilarity between visitors, users are quite hetergenous and we can pick 3 to 4 clusters that are stable. 
As we can see from plot of k against the wss measure, the slope of the graph changes in many situations, but we can consider that there is a significant drop of within sum of squared with k=4, the 5th weren't stable since the *wss* increases for the k=5.

# Using K-means with k=4

```{r, message=FALSE,warning=FALSE}

km=kmeans(Logstocluster.s,4,nstart=100,iter.max = 100)
summary(km)

Logstocluster$cluster=as.factor(km$cluster)
SmallLogs$cluster=as.factor(km$cluster)
```

```{r}
#Size of each cluster 
km$size
```


#Interpretation of clusters

```{r, message=FALSE,warning=FALSE,fig.height=7,fig.width=15}
#Plotting clusters against scenes 
library(ggplot2)
ggplot(data=subset(SmallLogs,scene_name%in%index),aes(x=scene_name,fill=cluster))+geom_bar(position = "fill")+theme(axis.text.x=element_text(angle=90))
```

```{r}
#Station
ggplot(data=SmallLogs,aes(x=station,fill=cluster))+geom_bar(position = "fill")+theme(axis.text.x=element_text(angle=90,size=9))
```

# Inference - KMeans Results 

The characterization of clusters is made by plotting variables used in the clustering and the resulting clusters. So the above plot help us to define the characteristic of each cluster:  

1.Dance, Example, Instruction, Idle, Game over corresponding to the first cluster and it correspond to the **cnk04** machine which is believed to be a gaming machine with dancing patterns.  

2.Stations **cnk02a** and **cnk02b** are believed to typical station machine, almost all users use these two stations who are believed to be regular stations at the museum that everyone uses.  

3.Station **cnk03** belong to the 4th cluster, and the most pattern used in this station are Multi level Main (LevelOneMain, LevelTwoMain,etc.)

The most used patterns are  more related to the stations **cnk02a** and **cnk02b**. As said before those patterns are Screen and Splash. 


##Partitioning Around Medoids

Partitioning around Medoids is robust version of k means. The similarity between the two  is due to mostly because to the fact that the two algorithms break the dataset into groups, and both work by trying to minimize the error, but PAM works with medoids, that are an entity of the dataset that represent the group in which it is inserted, and Kmeans works with centroids, that are artificially created entity that represent its cluster.

```{r}
library(cluster)
Logstocluster$time_inter=NULL
Logstocluster$visitor=NULL
max=nrow(Logstocluster)
set.seed(1223)
train_Logs=Logstocluster[sample(nrow(Logstocluster),nrow(Logstocluster)*0.01), ]
pam=pam(train_Logs[,-38],k=4)
pam$medoids

#to compare the k means and pam 
clusters_pam=unname(as.numeric(pam$clustering))
clusters_km=as.numeric(as.character(train_Logs[,38]))

#Percentage of similarity between the two clustering methods
diss=function(v,c){
  sum=0
  for(i in 1:length(c)){
    if (c[i]==v[i]) sum=sum+1
  }

return(sum)
  }
diss(clusters_km,clusters_pam)/nrow(train_Logs)
#0.35
```

The similarity between pam and K-means clustering is about 35%.

### Using PCA to confirm k-means clustering

We took the clustering from the k means Implementation and try to confirm it by principal component analysis. We plot the clusters in the two principal component of this reduction, and as we see, we've got almost 4 clusters. Clusters 1 and 4 are quite the same cluster.

```{r}
library(FactoMineR)
res=PCA(Logstocluster,quali.sup = 38,graph = FALSE)
plot(res,habillage=38)
plot(res,invisible="ind")
```

#Hierarchical clustering
```{r}

d=dist(train_Logs[,-max],method = "manhattan")
hc=hclust(d,method = "complete")
plot(hc,labels = train_Logs$cluster,cex=0.8)

```

# Conclusion 

It shows again three clusters 1,2 and 3. The  2nd and 3rd cluster are quite similar while the 1st cluster is a bit different.  
Comparing this method to k means confirmed the choice of k=4 and showed that the choice of three cluster is a good hint.
According to pam, there is a similarity in the way the two algorithms works with 35% of similarity. 
With no rule of thumb to choose the best clustering method, the k means and its interpretation are still valid to cluster the population of Logs into machines.

