---
title: "Phase2"
author: "Tomasz K, Tomasz Z, Margareta K"
date: "January 8, 2016"
output: html_document
---

#Introduction
In the second phase we will segment visitors into separate categories (segments) and answer the following questions:

 - how to define the similarity measure between visitors?
 - is the population homogeneous or heterogeneous? If heterogeneous then how many groups can we derive/define?
 - How to characterize different groups of visitors?
 - is there a pattern in stations that visitor tends to visit?

#Loading libraries and dataset
```{r, warnings = FALSE, cache = TRUE}
library(ggplot2)
library(MASS)
library(cluster)
library(tidyr)
library(sqldf)
library(reshape2)
library(cluster)
library(fpc)

# load("verySmallLogs.rda")
load("C:\\Users\\Tomek\\Desktop\\BISD\\Semestr 2\\Data mining\\Projekty\\4\\verySmallLogs.rda")
verySmallLogs$dayOfWeek = strftime(verySmallLogs$date,'%A')
verySmallLogs$dayOfWeekNumber = as.numeric(strftime(verySmallLogs$date,'%u'))
verySmallLogs$hour = substr(as.character(verySmallLogs$date), 12, 13);
verySmallLogs = verySmallLogs[order(verySmallLogs$dayOfWeekNumber), ]
verySmallLogs = verySmallLogs[!verySmallLogs$visitor == -1, ]
```

# Clustering

## How to define the similarity measure between visitors?

We decided to cluster visitors by the following measures:<br />
 1) days of week of visits<br />
 2) stations visited by visitors during one visit to the centre<br />
 3) how long did the visitor stay at any of the stations during one visit<br />
 4) number of times each station was visited by every visitor during one visit<br />

## Day of week clusters
```{r, warnings = FALSE, cache = TRUE}
visitor_day_of_week = sqldf('SELECT visitor, dayOfWeekNumber as cluster FROM verySmallLogs GROUP BY visitor, dayOfWeekNumber')
clustersCount = sqldf('SELECT cluster, count(*) FROM visitor_day_of_week group by cluster');
clustersCount
barplot(clustersCount$`count(*)`, main="How visitors fall into days of week", xlab="Days per week")
```
<br />
From the plot we can see that Friday is the most popular day amongst visitors, Saturday and Thursday following. Sunday is the least popular, as for Monday the station is closed. 

## Stations visited by visitors during one visit to the centre

###Matrix preparation
```{r, warnings = FALSE, cache=TRUE}
time_per_visitor <- sqldf('SELECT station, visitor, MAX(date) - MIN(date) as time FROM verySmallLogs GROUP BY station, visitor')

data <- time_per_visitor[,c(1,2)]
mat <- t(table(data))
mat <- mat[, colSums(mat != 0) > 0]
pmat = mat[1:5000, ]
```

###K-means
```{r, warnings = FALSE, cache=TRUE}
# run K-Means
km <- kmeans(mat, 3)

visitorsClustered = as.data.frame(mat);
visitorsClustered$cluster = km$cluster;
clustersCount = sqldf('SELECT cluster, count(*) FROM visitorsClustered group by cluster');
clustersCount

# Cluster plot
plotcluster(mat, km$cluster)
```

###PAM
```{r, warnings = FALSE, cache=TRUE}
pm <- pam(pmat, 3)

visitorsClustered = as.data.frame(pmat);
visitorsClustered$cluster = pm$cluster;
clustersCount = sqldf('SELECT cluster, count(*) FROM visitorsClustered group by cluster');
clustersCount

# Cluster plot
plotcluster(pmat, pm$cluster)
```

###Hierarchical clustering (binary)
```{r, warnings = FALSE, cache=TRUE}

# Hierarchical clustering (binary distance)
d = dist(pmat, method = "binary")
hc = hclust(d, method="ward.D")
# hc = agnes(d, method="ward")
plot(hc)

# Clusters
visitorsClustered = visitorsClustered2 = as.data.frame(pmat);
visitorsClustered$cluster = factor(cutree(hc, k=3));
clustersCount = sqldf('SELECT cluster, count(*) FROM visitorsClustered group by cluster');
clustersCount
plotcluster(pmat, visitorsClustered$cluster)
# ggplot(data=visitorsClustered,aes(x=cluster))+geom_bar(binwidth=1)
```

###Hierarchical clustering (euclidean)
```{r, warnings = FALSE, cache=TRUE}

# Hierarchical clustering (euclidean distance)
d = dist(pmat, method = "euclidean")
# Yo
hc = hclust(d, method="ward.D")
# hc = agnes(d, method="ward")
plot(hc)

# Clusters
visitorsClustered2$cluster = factor(cutree(hc, k=3));
clustersCount = sqldf('SELECT cluster, count(*) FROM visitorsClustered2 group by cluster');
clustersCount
plotcluster(pmat, visitorsClustered2$cluster)
```

## Clustering based on time spent at each station

```{r, warnings = FALSE, cache = TRUE}
# Time spent by visitors at each station
visitorsTimePerStation = dcast(time_per_visitor, visitor ~ station, value.var = "time", fill = 0)

# Scaling
visitorsTimePerStationMatrix <- as.matrix(scale(visitorsTimePerStation[, 2:ncol(visitorsTimePerStation)]))
``` 

###K-means
```{r, warnings = FALSE, cache = TRUE}

# Kmeans clustering by time
km <- kmeans(visitorsTimePerStationMatrix, 3)

# Clusters by time
visitorsClustered = as.data.frame(visitorsTimePerStationMatrix)
visitorsClustered$cluster = km$cluster
clustersCount = sqldf('SELECT cluster, count(*) FROM visitorsClustered GROUP BY cluster')
clustersCount

plotcluster(visitorsTimePerStationMatrix, km$cluster)
```

###PAM
```{r, warnings = FALSE, cache=TRUE}
pm <- pam(visitorsTimePerStationMatrix[1:5000,], 3)

visitorsClustered = as.data.frame(visitorsTimePerStationMatrix[1:5000,]);
visitorsClustered$cluster = pm$cluster;
clustersCount = sqldf('SELECT cluster, count(*) FROM visitorsClustered group by cluster');
clustersCount

# Cluster plot
plotcluster(mat[1:5000,], pm$cluster)
```

###Hierarchical clustering
```{r, warnings = FALSE, cache=TRUE}

distanceMatrix = dist(visitorsTimePerStationMatrix[1:5000,], method = "euclidean")
hc = hclust(distanceMatrix, method="ward.D")
plot(hc)

# Clusters by time
visitorsClustered = visitorsTimePerStation[1:5000, ];
visitorsClustered$cluster = factor(cutree(hc, k=3));
clustersCount = sqldf('SELECT cluster, count(*) FROM visitorsClustered group by cluster');
clustersCount
plotcluster(visitorsTimePerStationMatrix[1:5000,], visitorsClustered$cluster)
```

## Clustering based on number of times each station was visited
```{r, warnings = FALSE, cache=TRUE}

interactionsPerMachinePerVisit = sqldf('SELECT station, visitor, COUNT(*) as cnt FROM verySmallLogs WHERE type = "Entering" GROUP BY station, visitor');

interactionsPerMachinePerVisit = dcast(interactionsPerMachinePerVisit, visitor ~ station, value.var = "cnt", fill = 0)

# Scaling
interactionsPerMachinePerVisitMatrix <- as.matrix(scale(interactionsPerMachinePerVisit[, 2:ncol(interactionsPerMachinePerVisit)]))
```

###K-means
```{r, warnings = FALSE, cache=TRUE}
km <- kmeans(interactionsPerMachinePerVisitMatrix, 3)

# Clusters by time
visitorsClustered = as.data.frame(interactionsPerMachinePerVisitMatrix)
visitorsClustered$cluster = km$cluster
clustersCount = sqldf('SELECT cluster, count(*) FROM visitorsClustered GROUP BY cluster')
clustersCount

plotcluster(interactionsPerMachinePerVisitMatrix, km$cluster)
```

###PAM
```{r, warnings = FALSE, cache=TRUE}
pm <- pam(interactionsPerMachinePerVisitMatrix[1:5000,], 3)

visitorsClustered = as.data.frame(interactionsPerMachinePerVisitMatrix[1:5000,]);
visitorsClustered$cluster = pm$cluster;
clustersCount = sqldf('SELECT cluster, count(*) FROM visitorsClustered group by cluster');
clustersCount

# Cluster plot
plotcluster(interactionsPerMachinePerVisitMatrix[1:5000,], pm$cluster)
```

###Hierarchical clustering
```{r, warnings = FALSE, cache=TRUE}
distanceMatrix = dist(interactionsPerMachinePerVisitMatrix[1:5000,], method = "euclidean")
hc = hclust(distanceMatrix, method="ward.D")
plot(hc)

# Clusters by time
visitorsClustered = interactionsPerMachinePerVisit[1:5000,];
visitorsClustered$cluster = factor(cutree(hc, k=3));
clustersCount = sqldf('SELECT cluster, count(*) FROM visitorsClustered group by cluster');
clustersCount
plotcluster(interactionsPerMachinePerVisitMatrix[1:5000,], visitorsClustered$cluster)
```

#Summary

For using 3 clusters, we found out that the population is heterogeneous because we can clearly see 3 separate clusters when looking at plots generated using plotcluster on kmeans and pam clustering. There are 3 completely separated clusters in clustering based on stations visited by visitors during one visit to the centre with kmeans and pam method. Also there are 3 clusters in clustering based on number of times each station was visited and clustering based on time spent at each station using kmeans, however on these two similarity measures pam method didn't recognize distinct clusters which may be due to having used only 5000 visitors because of computing limitations.

When it comes to station utilization, we can describe the 3 groups in the following way. The first group of visitors uses most of the stations and does this also for a fairly long time (when compared with average station interaction time). Second group only briefly uses at most few stations for a relatively short period. Third group focuses on occupying few (1-3) stations for a relatively (very) long time. Visitors also differ by the number of times they play with machines. Some of them use certain machines (rather little) multiple times during one visit to the centre and don't play with other machines at all or just try them a bit. Other visitors try some machines for a smaller number of times. The last group tries most of the machines for a more or less equal (not very high) number of times.