---
title: "Homework"
author: "Marcel Sz"
date: "2015-10-14"
output: 
  html_document:
    toc: TRUE
---
#Homework description 
Use `knitr` to create a report for kNN method based on `GermanCredit` dataset.
Choose 5 variables and build classifier for them. The response variable is the 'Class'.

Find optimal `k` and calculate performance for it.

#Loading dataset and necessary library
```{r}
library(caret)
data(GermanCredit)
```
#Training classifier
```{r}
knnFit <- knn3(Class ~ Duration+Amount+Age+ResidenceDuration+NumberExistingCredits, data = GermanCredit, k=40, prob=TRUE)
knnFit
```

#Applying trained classifier (knnFit) to new data set
```{r}
pred <- predict(knnFit, data.frame(Amount = 1100, Duration = 24, NumberExistingCredits= 3, Age = 30, ResidenceDuration = 4))
t(pred)
par(mar=c(5,3,2,2))
barplot(pred, las=2)

# majority voting
pred <- predict(knnFit, data.frame(Amount = 1100, Duration = 24, NumberExistingCredits= 3, Age = 30, ResidenceDuration = 4), type="class")
```

#Constructing contingency table for the predicted labels to check quality of the classifier

```{r}
knnFit <- knn3(Class ~ Duration+Amount+Age+ResidenceDuration+NumberExistingCredits, data = GermanCredit, k=40)
pred <- predict(knnFit, GermanCredit, type="class")
pred
tab<-table(true = GermanCredit$Class, predicted = pred)
tab

sum(diag(tab)) / sum(tab)
```

#Dividing the dataset into two subsets: training and testing
```{r}
set.seed(1000)
indxTrain <- createDataPartition(y = GermanCredit$Class, p = 0.75)
str(indxTrain)

GermanCreditTrain <- GermanCredit[indxTrain$Resample1,]
GermanCreditTest <- GermanCredit[-indxTrain$Resample1,]
```

#Training classifier on the training dataset and testing it on the second dataset
```{r}
knnFit <- knn3(Class ~ Duration+Amount+Age+ CreditHistory.Delay+NumberExistingCredits, data = GermanCreditTrain, k=40, prob=TRUE)
knnFit
pred <- predict(knnFit, GermanCreditTest, type="class")

tab <- table(true = GermanCreditTest$Class, predicted = pred)
tab
tab2 <- prop.table(tab, 1)
tab2

sum(diag(tab)) / sum(tab)
sum(diag(tab2)) / sum(tab2)
```

#Choosing k and calculating performance

```{r}
tuneK <- 1:200

performance <- sapply(tuneK, function(k) {
  knnFit <- knn3(Class ~ Duration+Amount+Age+ResidenceDuration+NumberExistingCredits, data = GermanCreditTrain, k=k)
  tab <- table(true = GermanCreditTest$Class,
          predict = predict(knnFit, GermanCreditTest, type="class"))
  sum(diag(tab)) / sum(tab)
}) 

df <- data.frame(tuneK, performance)

ggplot(df, aes(tuneK, performance)) +
  geom_point() + 
  geom_smooth(se=FALSE, span=0.1, size=2) +
  theme_bw()
```