---
title: "ROC curves"
author: "Przemyslaw Biecek"
date: "Data Mining - Advances"
output: 
  html_document:
    toc: TRUE
---

# ROC (Receiver Operating Characteristics)

Classifiers that calculate scores / probabilities may be examined with the use of so called ROC. Let’s see an simple example for the trained random forest.

```{r, fig.width=8, fig.height=8}
library(randomForest)
library(faraway)
pima$test <- factor(pima$test)

ffit <- randomForest(test ~ .,   data=pima, importance = TRUE)
prob <- predict(ffit, type="prob")[,2]

library(ROCR)
fit.pred = prediction(prob, pima$test)
fit.perf = performance(fit.pred,"tpr","fpr")
plot(fit.perf)
abline(a=0,b=1)

fit.pred = prediction(prob, pima$test)
fit.perf = performance(fit.pred,"tpr","fpr")
plot(fit.perf, colorize=TRUE)
abline(a=0,b=1)

```

# Multiple curves

```{r, fig.width=8, fig.height=8}
library(caret)

ffit <- randomForest(test ~ .,   data=pima, importance = TRUE, mtry=3)
prob <- predict(ffit, type="prob")[,2]

ffit2 <- knn3(test ~ .,   data=pima, k=10)
prob2 <- predict(ffit2, newdata = pima, type="prob")[,2]

fit.pred = prediction(prob, pima$test)
fit.perf = performance(fit.pred,"tpr","fpr")
plot(fit.perf, col="red3")
fit.pred2 = prediction(prob2, pima$test)
fit.perf2 = performance(fit.pred2,"tpr","fpr")
plot(fit.perf2, col="blue3", add=TRUE)
abline(a=0,b=1)

```

Do this now on training and testing set!


# AUC (Area Under the Curve)

Functions from ROC package may work with other measures. See `?performance` for the full list.

```{r, fig.width=8, fig.height=8}
fit.pred = prediction(prob, pima$test)
fit.perf = performance(fit.pred,"auc")
fit.perf@y.values[[1]]
```


