---
title: "Hierarchical Clustering"
author: "Przemyslaw Biecek"
date: "Data Mining - Advances"
output: 
  html_document:
    toc: TRUE
---

# Clustering

Clustering is the task of grouping objects in such a way that objects in the same group (called a cluster) are more similar to each other than to those in other groups.

Clustering can be performed with the use of different methods/approaches. Today we are going to discuss hierarchical clustering.


There are two common approaches to build such hierarchy of clusters (also referred as a dendrogram):

- bottom up (agglomerative) - from small clusters to larger and larger,
- top down (divisive) - from large clusters to small ones.

First one is implemented in the function `diana{cluster}`, while the second one is implemented in the function `agnes{cluster}`. 

## The distance matrix

Let's use the dataset `Cars93` as an example. The dataset contains information about length and weight of different cars. 

```{r, fig.width=10, fig.height=10}
library(ggplot2)
library(MASS)
library(cluster)

ggplot(Cars93, aes(Length, Weight, label=Make)) +
  geom_text(size=3) + 
  theme_bw()
```

Can we find some groups in this dataset? Can we divide all cars into a few classes?

First, in order to find groups, we shall start with distances. How to calculate the matrix of distances?

It can be done with the `dist()` function. Different measures are implemented there. Typical choices:

- `euclidean`: usual distance between the two vectors $\sqrt(\sum((x_i - y_i)^2))$,
- `maximum`: maximum distance between two components of x and y,
- `manhattan`: absolute distance between the two vectors,
- `canberra`: $\sum(|x_i - y_i| / |x_i + y_i|)$ (usually used for counts),
- `binary`: vectors are regarded as binary bits, so non-zero elements are `on` and zero elements are `off`. The distance is the proportion of bits in which only one is on amongst those in which at least one is on.

But would it work in our case?

Not really. 

Note that for the two selected variables units are very very different. This is why we should scale these variables first. 

```{r}
# euclidian distances for original dataset
mat1 <- dist(Cars93[,c("Length","Weight")])
as.matrix(mat1)[1:5,1:5]

# euclidian distances for scaled dataset
mat2 <- dist(scale(Cars93[,c("Length","Weight")]))
as.matrix(mat2)[1:5,1:5]

# manhattan distances for scaled dataset
mat3 <- dist(scale(Cars93[,c("Length","Weight")]), method="manhattan")
as.matrix(mat3)[1:5,1:5]
```

## The algorithm

The clustering algorithm is a loop with two steps:

- Find two sub clusters that are closest to each other,
- Merge them into a single cluster until only one cluster remains.

## The linkage 

In the distance matrix one will find distances between particular observations. The linkage defines how to update distances after merging two clusters.

The commons choices are:

* `complete` 	 $\max \{\, d(a,b) : a \in A,\, b \in B \}$. 
* `single` 	$\min \{\, d(a,b) : a \in A,\, b \in B \}$. 
* `average`	 $\frac{1}{|A| |B|} \sum_{a \in A }\sum_{ b \in B} d(a,b)$. 
* `ward`, i.e. the minimum variance criterion (the preferred method).

Note that different methods lead to very different results. 

```{r}
rownames(Cars93) <- Cars93$Make
dat <- scale(Cars93[,c("Length","Weight")])

hc <- agnes(dat, method="complete")
plot(hc, which.plots=2, cex=0.5, main="")

hc <- agnes(dat, method="single")
plot(hc, which.plots=2, cex=0.5, main="")

hc <- agnes(dat, method="average")
plot(hc, which.plots=2, cex=0.5, main="")

hc <- agnes(dat, method="ward")
plot(hc, which.plots=2, cex=0.5, main="")
```

Agglomerative Coefficient is an average of distances between two consecutive nodes. The larger the better.

## Cutting the tree

Sometimes one would like to extract clusters from tree. 
You can use the `cutree` function for that.

How to choose the number of clusters? 
In most cases it's arbitrary choice. Sometimes this number can be extracted from the plot.
 
```{r}
hc <- agnes(dat, method="ward")
Cars93$labels = factor(cutree(hc, k=4))
ggplot(Cars93, aes(Length, Weight, label=Make, color=labels)) +
  geom_text(size=3) + 
  theme_bw()

hc <- agnes(dat, method="average")
Cars93$labels = factor(cutree(hc, k=4))
ggplot(Cars93, aes(Length, Weight, label=Make, color=labels)) +
  geom_text(size=3) + 
  theme_bw()

hc <- agnes(dat, method="single")
Cars93$labels = factor(cutree(hc, k=4))
ggplot(Cars93, aes(Length, Weight, label=Make, color=labels)) +
  geom_text(size=3) + 
  theme_bw()
```

## p-values for clusters

And sometimes you can use bootstrap probabilities (BP) / approximate unbiased probabilities (AU) for that.

```{r, message=FALSE, warning=FALSE, fig.width=10, fig.height=10}
library(pvclust)
dat <- scale(Cars93[Cars93$Origin == "non-USA",
                    c("Length","Weight","Width","Horsepower","Price")])
pv <- pvclust(t(dat))
plot(pv)
pvrect(pv, alpha=0.9)

```

## More plots

The `ggdendro` gives great types of plots.

```{r, fig.width=10, fig.height=10}
library(ape)
rownames(Cars93) <- Cars93$Make
dat <- scale(Cars93[,c("Length","Weight")])

library(RColorBrewer)
cols <- brewer.pal(3,"Set1")

hc <- as.phylo(as.hclust(agnes(dat, method="complete")))

par(mar=c(1,1,2,1), xpd=NA)

plot(hc, type = "fan", cex = 0.8,
     tip.color = cols[Cars93$Origin])

plot(as.phylo(hc), type = "unrooted", cex = 0.8,
     tip.color = cols[Cars93$Origin])

plot(as.phylo(hc), type = "radial", cex = 0.8,
     tip.color = cols[Cars93$Origin])

plot(as.phylo(hc), type = "phylogram", cex = 0.8,
     tip.color = cols[Cars93$Origin])

plot(as.phylo(hc), type = "cladogram", cex = 0.8,
     tip.color = cols[Cars93$Origin])
```


# Computer classes

Use data from votings of deputies from Polish Sejm (previous cadence).

The task: 
Cluster voting profiles for different deputies and present them graphically.


```{r}
load("all_votes.rda")
head(all_votes[,1:7])
```

Details:

- create a matrix with deputies in rows and id_voting in columns (you can use `spread()` function from `tidyr`),
- create matrix with similarities between deputies,
- use this similarity matrix to cluster deputies.

Try different distances and different methods for linkage. 

If you cannot create such matrix for all deputies, choose only deputies that were present during majority of votings.

Choose only deputies from two largest parties (PO and PiS).

Choose only important votings (these on which more than 75\% of deputies are present).

Check which deputies from party X have votes more similar to deputies from party B.

Show the dendrogram for selected deputies and use colours to present different parties.


