---
title: "Linear discriminant analysis (LDA) and Quadratic discriminant analysis (QDA)"
author: "Przemyslaw Biecek"
date: "Data Mining - Advances"
output: 
  html_document:
    toc: TRUE
---

# Linear discriminant analysis (LDA) 

This chapter is based on `An Introduction to Statistical Learning`
http://www-bcf.usc.edu/~gareth/ISL/ 
and chapter 4.3 from `The Elements of Statistical Learning` http://statweb.stanford.edu/~tibs/ElemStatLearn/

See also an overview here:
http://rstudio-pubs-static.s3.amazonaws.com/21386_bcf06f4cb32c4c408e282b906637ccdb.html

## S&P Stock Market Data

Daily percentage returns for the S&P 500 stock index between 2001 and 2005.

```{r}
library(ISLR)
head(Smarket)

library(ggplot2)
ggplot(Smarket, aes(Lag1, Lag2, color=Direction)) +
  geom_point()

ggplot(Smarket, aes(Lag1, Lag2, color=Direction)) +
  geom_density2d(h=1)

```

## LDA

Let's try a linear discrimination (https://en.wikipedia.org/wiki/Linear_discriminant_analysis).

Groups follow Gaussian distribution with same variance-covariance structure and different means.

```{r}
library(MASS)

lda.model <- lda(Direction~Lag1 + Lag2, data=Smarket)
lda.model

lda.pred <- predict(lda.model, Smarket)
names(lda.pred)

table(predicted = lda.pred$class, real = Smarket$Direction)
```

## Iris Data

Originally it was introduced for iris data.

```{r}
head(iris)

library(ggplot2)
ggplot(iris, aes(Sepal.Length, Sepal.Width, color=Species)) +
  geom_point()

ggplot(iris, aes(Sepal.Length, Sepal.Width, color=Species)) +
  geom_density2d(h=1)

```

# Quadratic discriminant analysis (QDA)

Just relax the assumption that variance-covariance matrix is the same in each group.

```{r}
qda.model <- qda(Direction ~ Lag1 + Lag2, data = Smarket)
qda.model


qda.class <- predict(qda.model, Smarket)$class
table(predicted = qda.class, real = Smarket$Direction)

```

# How to do this with caret

Here the pre-processing (scale/center) does not matter. But we show it just as an example how it can be done.

```{r}
library(caret)

modelFit<- train(Direction~Lag1+Lag2, method='lda',preProcess=c('scale', 'center'), data=Smarket)

confusionMatrix(Smarket$Direction, predict(modelFit, Smarket))
```

For both functions

```{r}

mat <- sapply(c('lda', 'qda'), function (met) {
  modelFit<- train(Direction~Lag1+Lag2, method=met,preProcess=c('scale', 'center'), data=Smarket)
  confusionMatrix(Smarket$Direction, predict(modelFit, Smarket))$overall
})

round(mat*100,2)
```

# The Homework

Use the k-fold cross validation to assess the performance of lda/qda on the wines dataset.

