---
title: 'DMA - Lab Task #1'
author: "Viet Ba Mai"
date: "11 października 2016"
output:
  html_document:
    toc: TRUE
---

#File Import

```{r, cache=TRUE}
allegro_trans <- read.csv(file="C:/Users/vietba/Documents/Computer Science/Masters/II/Data Mining Advances/Homeworks/Lab 1/allegro-api-transactions.csv", header=TRUE, sep=",", encoding="UTF-8")

allegro_cat <- read.table("C:/Users/vietba/Documents/Computer Science/Masters/II/Data Mining Advances/Homeworks/Lab 1/allegro-categories-eng.txt", header=TRUE, sep=";", quote=NULL, comment="", encoding="UTF-8")
```

##Tables
Below 10 first observations of each imported tables are presented.

1. Allegro Transactions
```{r, echo=FALSE, cache=TRUE}
head(allegro_trans, 10)
```

2. Allegro Categories (eng)
```{r, echo=FALSE, cache=TRUE}
head(allegro_cat, 10)
```


#Lookup Table and Translation
In the following case we will make use of the dplyr library.

##Joining the tables with left_join
```{r, cache=TRUE}
#install.packages("dplyr")

library(dplyr)
allegro_eng <- left_join(allegro_trans, allegro_cat, by = c("main_category" = "X.U.FEFF.main_category"))
```

##Merged tables result
To make the result clearer the number of columns selected was restricted to 3.

```{r, echo=FALSE, cache=TRUE}
head(allegro_eng[, c("item_id", "main_category", "main_category_eng")], 10)
```


#Summaries
In this section summaries of the following variables will be done:
- price

-it_is_allegro_standard

-it_is_brand_zone

-it_seller_rating

-date

We need to select top 5 categories hence we will find the frequencies of each category and then filter to only the best 5.

```{r, echo=FALSE, cache=TRUE}
sort(table(allegro_eng$main_category_eng), decreasing=TRUE)
```

Now we can see that the top 5 categories are:
1. Home & Garden
2. Clothing, Shoes, Accessories
3. Automotive
4. Babies
5. Beauty

And will create a new table with only these categories.

```{r, cache=TRUE}
top5 <- c(" Home & Garden", " Clothing, Shoes, Accessories", " Automotive", " Babies", " Beauty")
allegro_trans_top <- subset(allegro_eng, allegro_eng$main_category_eng %in% top5)
```

##Numerical Summaries
###Categorical Variables

1. Is Allegro Standard:

```{r, echo=FALSE, cache=TRUE}
sort(table(allegro_trans_top$it_is_allegro_standard), decreasing=TRUE)
```

2. Brand Zone:

```{r, echo=FALSE, cache=TRUE}
sort(table(allegro_trans_top$it_is_brand_zone), decreasing=TRUE)
```

3. Transaction Date:

```{r, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE, R.options=list(max.print=5)}
sort(table(allegro_trans_top$date), decreasing=TRUE)
```
The number of output was limited to 5.


###Continuous Variables

1. Price:

```{r, echo=FALSE, cache=TRUE}
summary(allegro_trans_top$price)
```

2. Seller Rating:

```{r, echo=FALSE, cache=TRUE}
summary(allegro_trans_top$it_seller_rating)
```

##Graphical Summaries
###Categorical Variables

1. Is Allegro Standard:

```{r, echo=FALSE, cache=TRUE}
all_standard <- table(allegro_trans_top$it_is_allegro_standard)
barplot(all_standard, horiz = FALSE, las=1, col="lightyellow")

```

2. Brand Zone:

```{r, echo=FALSE, cache=TRUE}
b_zone <- table(allegro_trans_top$it_is_brand_zone)
barplot(b_zone, las=1, col="khaki")

```


###Continuous Variables

1. Price:

```{r, echo=FALSE, cache=TRUE}
par(mar=c(5,5,2,2))
hist(allegro_trans_top$price, col="lightblue")
boxplot(allegro_trans_top$price, horizontal = TRUE)

```

2. Seller Rating:

```{r, echo=FALSE, cache=TRUE}
hist(allegro_trans_top$it_seller_rating, col="lightgreen")
boxplot(allegro_trans_top$it_seller_rating, horizontal = TRUE)
```

##2-variable representations
###Numerical Summaries for different discrete variables

1. Price by Brand Zone:


```{r, echo=FALSE, cache=TRUE}
by(allegro_trans_top$price, allegro_trans_top$it_is_brand_zone, summary)
```      


2. Seller Rating by Brand Zone:

```{r, echo=FALSE, cache=TRUE}
by(allegro_trans_top$it_seller_rating, allegro_trans_top$it_is_brand_zone, summary)
```

3. Price by Allegro Standard:


```{r, echo=FALSE, cache=TRUE}
by(allegro_trans_top$price, allegro_trans_top$it_is_brand_zone, summary)
```      


4. Seller Rating by Allegro Standard:

```{r, echo=FALSE, cache=TRUE}
by(allegro_trans_top$it_seller_rating, allegro_trans_top$it_is_brand_zone, summary)
```

##Graphical summary of continuous variables.

Seller Rating and Price:

```{r, echo=FALSE, cache=TRUE}
boxplot(allegro_trans_top$it_seller_rating, allegro_trans_top$price)
par(mar=c(5,15,2,2))

```

#Conclusions
Most of items are of Allegro standard and in Brand Zone 0.
On 2016-04-03 22:32:33 the most number of recorded transactions were made with the value being 105. Following that were transactions made around 2 hours earlier with the number being 95.

The mean price of items is 80.59 while the maximum found is 119000.00.
On average the seller's rating is 19020.
The mean prices are higher in zone 1, but the maximum price was recorded in the zone 0.
On the other hand the seller rating is greater in the zone 1.

Allegro standard seems to affect both the price and seller rating, because in both cases the aforementioned factors were higher.
