---
title: "10th homework"
author: "Margareta Kusan"
date: "December 26, 2015"
output: html_document
---

#With the use of all_votes data:

-for both PAM and k-means find clustering for different number of clusters (from 2 to 20). * for each observation calculate the silhouette and then calculate the average silhouette score.

-plot the average silhouette as a function of number of clusters.

Loading the libraries and dataset
```{r, warnings = FALSE}
library(ggplot2)
library(MASS)
library(cluster)
library(tidyr)
load("all_votes.rda")
head(all_votes[,1:7])
```

K-means
```{r, warning=FALSE, results='hide'}
data <- all_votes[,c(1,3,4)]
data$vote <- ifelse(data$vote == 'For', 1, ifelse(data$vote == 'Against', -1, 0))
head(data)
middle <- spread(data, key = id_voting, vote, fill=0)
row_names <- middle$surname_name
middle <- middle[,-1]

set.seed(4)
silhouettes = c()
i = 2
for (i in 2:20){
  model <- kmeans(middle, i, nstart=5)
  nd <- data.frame(model$centers)
  dissE <- daisy(as.matrix(middle)) 
  si <- silhouette(model$cl, dissE)
  silhouettes <- c(silhouettes, mean(si[,3]))
  plot(si, main="k-means")
}
```

PAM
```{r, warning=FALSE, results='hide'}
silhouettesp <- c()
i = 2
for (i in 2:20){
  modelp <- pam(middle, i)
  nd <- data.frame(modelp$medoids)
  sip <- silhouette(modelp)
  silhouettesp <- c(silhouettesp, mean(sip[,3]))
  plot(sip,main="PAM")
}

```

Average silhouette as a function of number of clusters
```{r}
x = c(2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,20)
plot(x, silhouettes, xlab="Number of clusters", ylab="Average silhouette", main="K-means")

plot(x, silhouettesp, xlab="Number of clusters", ylab="Average silhouette", main="PAM")
```
