---
title: "Homework 10"
author: "Ziad Al Bkhetan"
date: "January 6, 2016"
output: 
  html_document:
    toc : TRUE
---

#Introduction
In this Homework I will cluster the deputies depending on the voting, K-MEANS and PAM into from 2 to 20 clusters and calculate silhouette for each of them, then calculate the average silhouette then plot these values. 

# Data Loading
```{r , cache=FALSE, warning=FALSE, message=FALSE}
library(dplyr)
library(tidyr)
library(cluster)

load("D:/MSc Computer science and informatics/3rd semester/Data Mining - Advances/Homeworks/all_votes.rda")

```

# Data Preparation
we will use the records in the dataset, but we will change the features, so for each deputy we will check the vote type for each id_voting, and we will cluster them based on the id_voting values which are : 
  <br>* 1 : For
  <br>* 0 : Not present
  <br>* -1: Abstain
  <br>* -2: Against
<br>then we will get these clusters and calculate silhouette for them.

```{r , cache=FALSE, warning=FALSE, message=FALSE}
# Decode the votes 
all_votes_new = all_votes %>%
  mutate(
    voting_type = (ifelse(vote=='Against', -2, ifelse(vote=='For', 1, ifelse(vote=='Abstain', -1, 0 ) )))
  )

# prepare the final dataset
all_votes_new = all_votes_new[,c("surname_name","club", "id_voting","voting_type")]
final_data = all_votes_new %>%
  spread(id_voting, voting_type, fill=0)

# change the columns names, and prepare list of columns which will be used in clustering
names(final_data) <-make.names(names(final_data), unique = FALSE)
cols_names <- names(final_data)
cols_names <- cols_names[cols_names != "surname_name"]
cols_names <- cols_names[cols_names != "club"]

```

# K-Means Clustering
I will use K-means method woth cluster number from 2 to 20 and calculate the average silhouette for them, then I will plot them. 

```{r , cache=FALSE, warning=FALSE, message=FALSE}
k_meanslst <- list()
for (i in 2:21)
{
  k_meanslst[[i - 1]] <- kmeans(final_data[,cols_names], i)
}

des <- daisy(as.matrix(final_data[,cols_names]))
kmeans_res <- sapply(k_meanslst, function(x) mean(silhouette(x$cl, des)[, 3]))
plot(kmeans_res, main = "K-means", xlab = "clusters", ylab = "average silhouette", type = "o", xaxt="n", las = 1)

```

# K-means Results
Here we can see the silhouette results sorted for each clusters number:

```{r , cache=FALSE, warning=FALSE, message=FALSE}

cls_num <- list (2:21)
k_means_data <- as.data.frame(kmeans_res)
k_means_data <- cbind(k_means_data, cls_num)
k_means_data = arrange(k_means_data, desc(kmeans_res))
for (i in 2:21)
{
  print (paste("Clusters number: ", k_means_data[i - 1, 2], ", Average silhouette:", k_means_data[i - 1, 1]))  
}
```

# PAM Clustering
I will do the same as in the prevoius part but using PAM instead of k-means, then I will plot the results
```{r , cache=FALSE, warning=FALSE, message=FALSE}

pam_lst <- list()
for (i in 2:21)
{
  pam_lst[[i - 1]] <- pam(final_data[,cols_names], i)
}

pam_res <- sapply(pam_lst, function(x) mean(silhouette(x)[, 3]))
plot(pam_res, main = "Pam", xlab = "clusters", ylab = "average silhouette", type = "o", xaxt="n", las = 1)


```

# Pam Results
Here we can see the silhouette results sorted for each clusters number:

```{r , cache=FALSE, warning=FALSE, message=FALSE}
Pam_data <- as.data.frame(pam_res)
Pam_data <- cbind(Pam_data, cls_num)
Pam_data = arrange(Pam_data, desc(pam_res))
for (i in 2:21)
{
  print (paste("Clusters number: ", Pam_data[i - 1, 2], ", Average silhouette:", Pam_data[i - 1, 1]))  
}

```



# Remarks
in tis homework we tried to compare the clustering with different clusters number, and using two different methods: k-means and pam.
<br> this comparision based on silhouette values, and we can find that the best value for k-means was :
<br>Clusters number:  3 , silhouette: 0.411187745669155
<br> and for Pam is:
<br>Clusters number:  2 , silhouette: 0.437521905746973
