---
title: "Homework 4"
author: "Karolina Kwasiborska"
date: "29-10-2015"
output: html_document
---

# The Homework

Download the dataset `adult` (find more about this dataset here: http://archive.ics.uci.edu/ml/machine-learning-databases/adult/). 

Train a k-nn for different k and Random Forest for different `m` and compare these results with the use of ROC curves and AUC values. 

```{r}
data = read.table("http://archive.ics.uci.edu/ml/machine-learning-databases/adult/adult.data",
        sep=",",header=F,col.names=c("age", "type_employer", "fnlwgt", "education", 
                "education_num","marital", "occupation", "relationship", "race","sex",
                "capital_gain", "capital_loss", "hr_per_week","country", "income"),
        fill=FALSE,strip.white=T)


data$income = factor(data$income)

# Training and testing sets
library(caret)
indxTrain <- createDataPartition(y = data$income, p = 0.75)
str(indxTrain)

trainSet <- data[indxTrain$Resample1,]
testSet <- data[-indxTrain$Resample1,]

```


# ROC (Receiver Operating Characteristics)

## Random forests
```{r, fig.width=8, fig.height=8}
library(randomForest)

ffit1 <- randomForest(income ~ .,   data=trainSet, importance = TRUE, mtry=3)
prob1 <- predict(ffit1, testSet, type="prob")[,2]

ffit2 <- randomForest(income ~ .,   data=trainSet, importance = TRUE, mtry=5)
prob2 <- predict(ffit2, testSet, type="prob")[,2]

ffit3 <- randomForest(income ~ .,   data=trainSet, importance = TRUE, mtry=10)
prob3 <- predict(ffit3, testSet, type="prob")[,2]

library(ROCR)

fit.pred = prediction(prob1, testSet$income)
fit.perf = performance(fit.pred,"tpr","fpr")
plot(fit.perf, col="red3")

fit.pred2 = prediction(prob2, testSet$income)
fit.perf2 = performance(fit.pred2,"tpr","fpr")
plot(fit.perf2, col="blue3", add=TRUE)

fit.pred3 = prediction(prob3, testSet$income)
fit.perf3 = performance(fit.pred3,"tpr","fpr")
plot(fit.perf3, col="black", add=TRUE)

```

## k-nn
```{r, fig.width=8, fig.height=8}

ffit4 <- knn3(income ~ .,   data=trainSet, k=5)
prob4 <- predict(ffit4, newdata = testSet, type="prob")[,2]

ffit5 <- knn3(income ~ .,   data=trainSet, k=15)
prob5 <- predict(ffit5, newdata = testSet, type="prob")[,2]

ffit6 <- knn3(income ~ .,   data=trainSet, k=50)
prob6 <- predict(ffit6, newdata = testSet, type="prob")[,2]

fit.pred4 = prediction(prob4, testSet$income)
fit.perf4 = performance(fit.pred4,"tpr","fpr")
plot(fit.perf4, col="green3")

fit.pred5 = prediction(prob5, testSet$income)
fit.perf5 = performance(fit.pred5,"tpr","fpr")
plot(fit.perf5, col="orange3", add=TRUE)

fit.pred6 = prediction(prob6, testSet$income)
fit.perf6 = performance(fit.pred6,"tpr","fpr")
plot(fit.perf6, col="yellow3", add=TRUE)
abline(a=0,b=1)

```

# AUC (Area Under the Curve)

## Random forests
```{r, fig.width=8, fig.height=8}
#Random forest mtry=3
fit.pred1 = prediction(prob1, testSet$income)
fit.perf1 = performance(fit.pred1,"auc")
fit.perf1@y.values[[1]]

#Random forest mtry=5
fit.pred2 = prediction(prob2, testSet$income)
fit.perf2 = performance(fit.pred2,"auc")
fit.perf2@y.values[[1]]

#Random forest mtry=10
fit.pred3 = prediction(prob3, testSet$income)
fit.perf3 = performance(fit.pred3,"auc")
fit.perf3@y.values[[1]]
```

## k-nn
```{r, fig.width=8, fig.height=8}
#knn k=5
fit.pred4 = prediction(prob4, testSet$income)
fit.perf4 = performance(fit.pred4,"auc")
fit.perf4@y.values[[1]]

#knn k=20
fit.pred5 = prediction(prob5, testSet$income)
fit.perf5 = performance(fit.pred5,"auc")
fit.perf5@y.values[[1]]

#knn k=50
fit.pred6 = prediction(prob6, testSet$income)
fit.perf6 = performance(fit.pred6,"auc")
fit.perf6@y.values[[1]]
?randomForest
```

# Summary

## Random forests
From both ROC curve and AuC we can see that the best result gives a random forest with mtry=3, the worst a random forest with mtry=10.

## knn
From both ROC curve and AuC we can see that the best result gives k-nn with k=5, the worst k-nn with k=50.
