---
title: "Homework 4"
author: "Marcel Sz"
date: "23 paüdziernika 2015"
output: html_document
---

# Homework 4
## Homework description 

Download the dataset adult (find more about this dataset here: http://archive.ics.uci.edu/ml/machine-learning-databases/adult/).
Train a k-nn for different k and Random Forest for different m and compare these results with the use of ROC curves and AUC values. 

##Loading libraries and data

```{r}
library(faraway)
library(MASS)
library(ggplot2)
library(randomForest)
library(caret)
library(ROCR)
data = read.table("http://archive.ics.uci.edu/ml/machine-learning-databases/adult/adult.data",
        sep=",",header=F,col.names=c("age", "type_employer", "fnlwgt", "education", 
                "education_num","marital", "occupation", "relationship", "race","sex",
                "capital_gain", "capital_loss", "hr_per_week","country", "income"),
        fill=FALSE,strip.white=T)
head(data,3)
```

# Knn for k=4 and Random Forest for m=2
```{r}
ffit <- randomForest(sex ~ education_num + age + capital_gain, data=data, importance = TRUE, mtry=2)
prob <- predict(ffit, type="prob")[,2]

ffit2 <- knn3(sex ~ education_num + age + capital_gain, data=data, k=4)
prob2 <- predict(ffit2, newdata = data, type="prob")[,2]

fit.pred = prediction(prob, data$sex)
fit.perf = performance(fit.pred,"tpr","fpr")
plot(fit.perf, col="yellow")

fit.pred2 = prediction(prob2, data$sex)
fit.perf2 = performance(fit.pred2,"tpr","fpr")
plot(fit.perf2, col="green", add=TRUE)
abline(a=0,b=1)
```

## Random Forest: AUC 

```{r}
fit.pred = prediction(prob, data$sex)
fit.perf = performance(fit.pred,"auc")
fit.perf@y.values[[1]]
```

## Knn: AUC

```{r}
fit.pred = prediction(prob2, data$sex)
fit.perf = performance(fit.pred,"auc")
fit.perf@y.values[[1]]
```

# Knn for k=8 and Random Forest for m=3
```{r}
ffit <- randomForest(sex ~ education_num + age + capital_gain,   data=data, importance = TRUE, mtry=3)
prob <- predict(ffit, type="prob")[,2]

ffit2 <- knn3(sex ~ education_num + age + capital_gain, data=data, k=8)
prob2 <- predict(ffit2, newdata = data, type="prob")[,2]

fit.pred = prediction(prob, data$sex)
fit.perf = performance(fit.pred,"tpr","fpr")
plot(fit.perf, col="yellow")

fit.pred2 = prediction(prob2, data$sex)
fit.perf2 = performance(fit.pred2,"tpr","fpr")
plot(fit.perf2, col="green", add=TRUE)
abline(a=0,b=1)
```

# Random Forest: AUC 

```{r}
fit.pred = prediction(prob, data$sex)
fit.perf = performance(fit.pred,"auc")
fit.perf@y.values[[1]]
```

# Knn: AUC

```{r}
fit.pred = prediction(prob2, data$sex)
fit.perf = performance(fit.pred,"auc")
fit.perf@y.values[[1]]
```
