---
title: "Homework 4"
author: "Ziad Al Bkhetan"
date: "October 29, 2015"
output: 
  html_document:
    toc : TRUE
---


# Data Loading

```{r}
adult_data = read.table("http://archive.ics.uci.edu/ml/machine-learning-databases/adult/adult.data",
        sep=",",header=F,col.names=c("age", "type_employer", "fnlwgt", "education", 
                "education_num","marital", "occupation", "relationship", "race","sex",
                "capital_gain", "capital_loss", "hr_per_week","country", "income"),
        fill=FALSE,strip.white=T)
head(adult_data,3)
```

# Data Partitioning
in this phase we will divide the dataset into two different sets the first one for classifier training while the second is for the classifier testing.
<br> * partion percentage is 75:25
<br> * target attribute to maintain good distribution "income"
 
```{r}
library(lattice)
library(ggplot2)
library(caret)
indxTrainSet <- createDataPartition(y = adult_data$income, p=0.75)
str(indxTrainSet)

adultDataTrain <- adult_data[indxTrainSet$Resample1,]
adultDataTest <- adult_data[-indxTrainSet$Resample1,]

```

# Libraries loading

```{r}
library(randomForest)
library(caret)
library (gplots)
library(ROCR)

```

# ROC Curve for Random Forersts and K Nearest Neighbors

ROC curves for Random forests are in blue 
<br>
ROC curves for K nearest Neighbors are in red

k, m from 1 to 5, you can change them in the script, just change max_counter variable

```{r}
max_counter = 5
rf <-c()
knn <-c()
for(i in 1:max_counter) 
{
  
  rfTree <- randomForest(income ~ capital_gain+capital_loss+marital+fnlwgt ,na.action = na.omit, data=adultDataTrain, importance=TRUE, mtry=i)
  rfTab <- predict(rfTree, newdata=adultDataTest, type="prob")[,2]
  
  fit.pred <- prediction(rfTab, adultDataTest$income)
  fit.perf <- performance(fit.pred,"tpr","fpr")
  if (i < 2)
  {
    plot(fit.perf, col="blue3")  
  }else
  {
    plot(fit.perf, col="blue3", add=TRUE)
  }
  
  fit.perf <- performance(fit.pred,"auc")
  rf<- c(rf, fit.perf@y.values[[1]])
  
  # KNN Classifier
  knnc <- knn3(income ~ capital_gain+capital_loss+marital+fnlwgt ,   data=adultDataTrain, k=i)
  kntab <- predict(knnc, newdata = adultDataTest, type="prob")[,2]
  
  fit.pred <- prediction(kntab, adultDataTest$income)
  fit.perf <- performance(fit.pred,"tpr","fpr")
  plot(fit.perf, col="red3", add=TRUE)
  
  fit.perf <- performance(fit.pred,"auc")
  knn<- c(knn, fit.perf@y.values[[1]])
  

  
}

abline(a=0,b=1)


```

# AUC values

for Random forests:
```{r}
for (j in 1:max_counter)
{
  cat("m = " , j, "  auc = ", rf[j], "\n")
  
}

```


for K Nearest Neighbors :
```{r}
for (j in 1:max_counter)
{
  cat("k = " , j, "  auc = ", knn[j], "\n")
}

```

