---
title: "Homework 5"
author: "Tomasz K"
date: "5 listopada 2015"
output: html_document
---

### Loading data
```{r, cache=TRUE, warning=FALSE, message=FALSE}
library(caret)
wines <- read.table("http://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-red.csv", sep=";", header=TRUE)
table(wines$quality)

winesb <- wines
winesb$quality <- factor(ifelse(wines$quality > 5, "good", "bad")) 
```

### Dividing variables into categories of equal length and size
```{r, cache=TRUE, warning=FALSE, message=FALSE}
winesb_length <- winesb
for (i in 1:11) {
  winesb_length[,i] <- cut(winesb[,i], 3)
}

winesb_size <- winesb
for (i in 1:11) {
  winesb_size[,i] <- cut_number(winesb[,i], 3)
}
```

### Calculating performance for both divisions
```{r, cache=TRUE, warning=FALSE, message=FALSE}
train_control <- trainControl(method="boot", number=10)

model_length <- train(quality~., data=winesb_length, trControl=train_control, method="nb")
predictions <- predict(model_length, winesb_length)
confusionMatrix(predictions, winesb_length$quality)

model_size <- train(quality~., data=winesb_size, trControl=train_control, method="nb")
predictions <- predict(model_size, winesb_size)
confusionMatrix(predictions, winesb_size$quality)
```

We can see that with categories of equal size the performance is better (accuracy: 0.7104 vs 0.6554; kappa: 0.4199 vs 0.2816).

### Checking feature importance
```{r, cache=TRUE, warning=FALSE, message=FALSE}
varImp(model_length, scale=TRUE)
plot(varImp(model_length))

varImp(model_size, scale=TRUE)
plot(varImp(model_size))
```

The three best features in case of categories of equal length are therefore: alcohol, citric.acid, fixed.acidity.

The three best features in case of categories of equal size are therefore: alcohol, sulphates, citric.acid.

### Calculating performance for both divisions using only three best features
```{r, cache=TRUE, warning=FALSE, message=FALSE}
train_control <- trainControl(method="boot", number=10)

model <- train(quality~alcohol+citric.acid+fixed.acidity, data=winesb_length, trControl=train_control, method="nb")
predictions <- predict(model, winesb_length)
confusionMatrix(predictions, winesb_length$quality)

model <- train(quality~alcohol+sulphates+citric.acid, data=winesb_size, trControl=train_control, method="nb")
predictions <- predict(model, winesb_size)
confusionMatrix(predictions, winesb_size$quality)
```

In case of using three best features with categories of equal length the performance is better (accuracy: 0.6948 vs 0.6554; kappa: 0.3969 vs 0.2816).

In case of using three best features with categories of equal size the performance is worse (accuracy: 0.6829 vs 0.7104; kappa: 0.3753 vs 0.4199).
