---
title: 'DMA - Homework #1'
author: "Viet Ba Mai"
date: "10 paüdziernika 2016"
output:
  html_document:
    toc: TRUE
---

#Description
In this homework we used the PogromcyDanych package along with its function setLang() which translates a dataset between English and Polish depending and the given parameter.
The goal was to create summaries, both numerical and graphical, for the dataset skiJumps2013 which represent real data from ski jumping competitions held in 2013 and 2014. 
The summaries were done on the following variables:

- jumpSpeed

- jumpDistance

- compCountry

- jumperCountry

- jumperDateOfBirth.

This report is done only on data where the competition was in 2014.


10 observations of selected columns from the skiJumps2013 table (with a date filter applied on) are presented below.

```{r, include=FALSE}
  # 1. Load the PogromcyDanych package.
  # 2. Use the setLang() function to translate columns names into English.
  # 3. For the dataset skiJumps2013 present numerical and graphical summaries for following variables: jumpSpeed,   # jumpDistance, compCountry, jumperCountry, jumperDateOfBirth.
  # 4. Create a report in knitr with these summaries. Use only competitions in 2014 year.
#install.packages("PogromcyDanych")
library(PogromcyDanych)
setLang("eng")
data(skiJumps2013)
skiJumps2014 <- subset(skiJumps2013, as.Date(skiJumps2013$compDate) >= "2014-01-01")
```

```{r, echo=FALSE, cache=TRUE}
head(skiJumps2014[, c("compDate", "jumpSpeed", "jumpDistance", "compCountry", "jumperCountry", "jumperDateOfBirth")], 10)
#head(skiJumps2014, 10)
```

#Numerical Summaries
##Categorical Variables
For categorical values we use contingency tables. This is done by using the table() function.

1. Competition Country:

```{r, echo=FALSE, cache=TRUE}
sort(table(skiJumps2014$compCountry), decreasing=TRUE)
```

2. Jumper's Country:

```{r, echo=FALSE, cache=TRUE}
sort(table(skiJumps2014$jumperCountry), decreasing=TRUE)
```

3. Jumper's Date Of Birth:

```{r, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE, R.options=list(max.print=15)}
sort(table(skiJumps2014$jumperDateOfBirth), decreasing=TRUE)
```
The number of output was limited to 15.


##Continuous Variables
For continuous values we use the summary() function.

1. Jump Speed:

```{r, echo=FALSE, cache=TRUE}
summary(skiJumps2014$jumpSpeed)
```

2. Jump Distance:

```{r, echo=FALSE, cache=TRUE}
summary(skiJumps2014$jumpDistance)
```

#Graphical Summaries
##Categorical Variables
Barplots [barplot()] are very well suited to visualize contingency tables.

1. Competition Country:

```{r, echo=FALSE, cache=TRUE}
comp_country <- table(skiJumps2014$compCountry)
barplot(comp_country, horiz = TRUE, las=1, col="lightyellow")

```

2. Jumper's Country:

```{r, echo=FALSE, cache=TRUE}
jumper_country <- table(skiJumps2014$jumperCountry)
barplot(jumper_country, las=2, col="khaki")

```

3. Jumper's Year Of Birth:

```{r, echo=FALSE, cache=TRUE}
jumper_bday <- table(as.numeric(format(skiJumps2014$jumperDateOfBirth, '%Y', col="lightpurple")))
barplot(jumper_bday, las=2)
```

Years were extracted from full dates to make the graph clearer.


##Continuous Variables
Histograms [hist()] and boxplots [boxplot()] can be used to represent the distribution of continuous values.


1. Jump Speed:

```{r, echo=FALSE, cache=TRUE}
par(mar=c(5,5,2,2))
hist(skiJumps2014$jumpSpeed, col="lightblue")
boxplot(skiJumps2014$jumpSpeed, horizontal = TRUE)

```

2. Jump Distance:

```{r, echo=FALSE, cache=TRUE}
hist(skiJumps2014$jumpDistance, col="lightgreen")
boxplot(skiJumps2014$jumpDistance, horizontal = TRUE)
```

#2-variable representations
##Numerical Summaries for different discrete variables

1. Jump Speed by Jumper Country:


```{r, echo=FALSE, cache=TRUE}
by(skiJumps2014$jumpSpeed, skiJumps2014$jumperCountry, summary)
```      


2. Jump Distance by Jumper Country:

```{r, echo=FALSE, cache=TRUE}
by(skiJumps2014$jumpDistance, skiJumps2014$jumperCountry, summary)
```

##Graphical summary for two discrete variables
Mosaicplots may be useful to show relations between different categorical variables.

Competition Country and Jumper Country:

```{r, echo=FALSE, cache=TRUE}
countries <- table(skiJumps2014$compCountry, skiJumps2014$jumperCountry)
#countries <- table(skiJumps2014$compCountry, as.numeric(format(skiJumps2014$jumperDateOfBirth)))
mosaicplot(countries, las=2, main="", color = c("orange", "yellow"))
```

##Numerical summary of relation between discrete and continuous variables.

Jumper Country and Jump Distance:

```{r, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE, R.options=list(max.print=25)}
country_bday <- table(skiJumps2014$jumperCountry, skiJumps2014$jumpDistance)
round(prop.table(country_bday, 2),1)
```
The number of output was limited to 25.


##Graphical summary of continuous variables.
Again, boxplots are very good for this type of variables. Additionally, a more advanced way is to use vioplot [vioplot()].

Jump Distance and Jump Speed:

```{r, echo=FALSE, cache=TRUE}
boxplot(skiJumps2014$jumpDistance, skiJumps2014$jumpSpeed)
par(mar=c(5,15,2,2))
#install.packages("vioplot")
#library(vioplot)
#vioplot(skiJumps2014$jumperCountry[skiJumps2014$jumpSpeed < 75],
#        skiJumps2014$jumperCountry[skiJumps2014$jumpSpeed >= 75])

```

#Conclusions
Most of competitions were held in, respectively, Austria, Germany and Finland. The first two countries also have the biggest number of jumpers.

Most of the competitors were born in 1991 and the second most frequent birth year is 1989 with the frequency being almost only a half of the aforementioned year.

In terms of the jump itself, the usual speed is around 80 to 100 while the distance is between 100 and 150.

Germany has the highest maximum jumping speed (101.50) while Slovenia has the highest maximum jumping distance (199.0).
On the other hand Romania has the lowest mean jumping speed (88.97) and Japan has the lowest mean in jumping distance (112.0).